### 2020-05-30 paper/paper.tex

\documentclass[11pt,english,twocolumn]{article}

\title{Append-only Datastore}
\author{
	Mingyan Zhao
	\and
	Steven Tung
	\and
	Kevin Krakauer
}
\date{}

\begin{document}

\maketitle

% Calling our operation "put" is inherently confusing, I'm using "append".

\begin{abstract}
Append-only datastore is an eventually consistent, append-only datastore focused
on low read/write latency and high availability. In the common case, clients
interact with nearby nodes only for extremely low latency. We store every input
value for a given key in a value list, but do not provide overwrite and delete
operations. We provide a simple interface that guarantees properties making
client implementation simple: the relative order of data for a given key is
consistent and only requires communication with a single node.

The datastore is eventually consistent. Data is stored in memory for low latency
and in local disk for safety. Nodes can operate when disconnected from the
system, preserving availability in the face of total and long-term partitioning.
We believe this system will be useful in chat and distributed logging
applications.
\end{abstract}

\section{Introduction}
La-de-da-de-da an intro.

\section{Design}
We made it go super fast like vroom-vroom.

\subsection{Leader}
It synchronizes n'stuff.

\subsection{Follower}
% Be sure to discuss any synchronization that happens here, as the follower is
% responsible for synchronizing its the writes to it.
It also sorta synchronizes, but mostly it follows.

\subsection{Clients}
% Describe the API and guarantees exposed to clients.
We have very distinguished clientele.

\section{Implementation}

\section{Evaluation}
We tested using some super cool gcloud stuff.

\subsection{Workload 1}
Oh wow, we're so fast.

\subsection{Workload 2}
Oh no, we're unexpectedly not that fast.

\section{Related Work}
This is the most filler of all filler sections. Look, a reference! \cite{dynamo}

\section{Future Work}
Make it actually work.

\section{Conclusions}
I'd just like everyone to pat themselves on the back here.

\bibliography{paper} 
\bibliographystyle{ieeetr}

\end{document}

### 2020-05-30 client/client.go
Implement simple broadcasting

@@ -77,7 +77,6 @@ func getData(c flpb.FollowerClient) {
 		if err != nil {
 			log.Fatalf("%v.Get(_) = _, %v: ", c, err)
 		}
-		log.Printf("Get(%v) =  %v", k, resp)
 		result = append(result, resp.Values...)
 	}
 	sort.Strings(result)
@@ -114,10 +113,14 @@ func main() {
 	// Send requests
 	sentRequests(client)
 
+	// Get the data
+	log.Printf("Get data first time")
+	getData(client)
+
 	// Wait for other test clients to finish
-	time.Sleep(3 * time.Second)
-	log.Printf("Clinet done")
+	time.Sleep(10 * time.Second)
 
 	// Get the data
+	log.Printf("Get data second time")
 	getData(client)
 }

### 2020-05-30 follower/follower.go

@@ -67,7 +67,8 @@ type data struct {
 	updateReqChan chan *lpb.UpdateRequest
 
 	// syncReqChan is the channel that sending the sync requests to the target followers
-	syncReqChan chan *syncRequest
+	syncReqChan  chan *syncRequest
+	syncReqQueue []*syncRequest
 
 	// syncRespChan is the channel that handles the received sync responses.
 	syncRespChan chan *pb.SyncResponse
@@ -133,6 +134,7 @@ func (f *follower) Put(ctx context.Context, req *pb.PutRequest) (*pb.PutResponse
 		f.store[req.Key] = &data{
 			updateReqChan:   make(chan *lpb.UpdateRequest, 100),
 			syncReqChan:     make(chan *syncRequest, 100),
+			syncReqQueue:    make([]*syncRequest, 0),
 			done:            make(chan bool),
 			buffer:          make([]string, 0),
 			versionToValues: make(map[int64][]string),
@@ -175,8 +177,9 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 			result = append(result, v...)
 
 			// then append all of the data that still waiting for update.
-			result = append(result, f.store[req.Key].buffer...)
+
 		}
+		result = append(result, f.store[req.Key].buffer...)
 	}
 
 	log.Printf("Get result is: [%v]", result)
@@ -239,6 +242,10 @@ func (f *follower) sendUpdate(key string) {
 	f.store[key].updateReqChan <- updateReq
 }
 
+func (f *follower) sendSync(data *data, req *syncRequest) {
+	data.syncReqQueue = append(data.syncReqQueue, req)
+}
+
 func (f *follower) handleSync(key string) error {
 	log.Printf("start sync handler for key %v", key)
 	syncReqChan := f.store[key].syncReqChan
@@ -279,16 +286,18 @@ func (f *follower) handleSync(key string) error {
 				if d, ok := myData[v.Version]; ok {
 					err := fmt.Errorf("something went wrong, the version number should not be conflicting, key %v, existing data %v, incomming data %v", key, d, v)
 					log.Println(err)
-					return err
+					continue
 				}
 				myData[v.Version] = v.Value
 			}
 			f.store[key].versionToValueLock.Unlock()
+
 			log.Printf("SyncResp done, %v", syncResp)
 		}
 	}
 }
 
+// askForVersions generates a list of versions in an open range of (localVer, globalVer)
 func askForVersions(globalVer, localVer int64) *pb.AskFor {
[…]

### 2020-05-30 leader/leader.go

@@ -8,11 +8,13 @@ import (
 	"log"
 	"net"
 	"sync"
+	"time"
 
 	"google.golang.org/grpc"
 
 	config "distributed-key-value-store/config"
 	cpb "distributed-key-value-store/protos/config"
+	fpb "distributed-key-value-store/protos/follower"
 	pb "distributed-key-value-store/protos/leader"
 	"distributed-key-value-store/util"
 )
@@ -22,8 +24,9 @@ const (
 	casterFlag          = "caster"
 	casterFlagImmediate = "immediate"
 	casterFlagPeriodic  = "periodic"
-
-	timeoutFlag = "timeout"
+	// Simple braodcasting means that the leader send a Notification of all of the key-version info to all of followers
+	casterFlagSimple = "simple"
+	timeoutFlag      = "timeout"
 
 	keyVersionMapConcurrency = 32
 )
@@ -67,6 +70,10 @@ type leader struct {
 
 	// broadcaster sends updates to all followers when an update occurs.
 	broadcaster broadcaster
+
+	done chan bool
+
+	followers map[string]fpb.FollowerClient
 }
 
 type versioninfo struct {
@@ -86,6 +93,7 @@ func newLeader(configuration *cpb.Configuration) (*leader, error) {
 	// Pick a broadcast mechanism.
 	var bc broadcaster
 	var err error
+	var follwers map[string]fpb.FollowerClient
 	switch *caster {
 	case casterFlagImmediate:
 		if bc, err = newImmediate(configuration); err != nil {
@@ -93,6 +101,11 @@ func newLeader(configuration *cpb.Configuration) (*leader, error) {
 		}
 	case casterFlagPeriodic:
 		if bc, err = newPeriodic(configuration); err != nil {
+			return nil, err
+		}
+	case casterFlagSimple:
+		follwers, err = loadFollowers(configuration)
+		if err != nil {
 			return nil, err
 		}
 	default:
@@ -103,6 +116,7 @@ func newLeader(configuration *cpb.Configuration) (*leader, error) {
 		configuration: configuration,
 		keyVersionMap: keyvaluemap{data: make(map[string]*versioninfo)},
 		broadcaster:   bc,
+		followers:     follwers,
 	}, nil
 }
 
@@ -169,6 +183,55 @@ func (l *leader) Update(ctx context.Context, req *pb.UpdateRequest) (*pb.UpdateR
 	return resp, nil
 }
 
[…]

### 2020-05-31 paper/paper.tex
Write some of the paper.
- Intro written
  - Background
  - System advantages
  - Use cases
- There's a 'make spell'

@@
 
-\begin{abstract}
-Append-only datastore is an eventually consistent, append-only datastore focused
-on low read/write latency and high availability. In the common case, clients
-interact with nearby nodes only for extremely low latency. We store every input
-value for a given key in a value list, but do not provide overwrite and delete
-operations. We provide a simple interface that guarantees properties making
-client implementation simple: the relative order of data for a given key is
-consistent and only requires communication with a single node.
+\section*{}
+We built an eventually consistent, append-only datastore focused on low
+read/write latency and high availability. In the common case, clients interact
+only with nearby nodes for extremely low latency. We do not provide overwrite
+and delete operations. This simple interface guarantees properties making client
+implementation simple: the relative order of data for a given key is consistent
+and each client communicates only with a single node.
 
@@
 applications.
-\end{abstract}
 
 \section{Introduction}
-La-de-da-de-da an intro.
+As organizations increasingly move to the cloud, applications are designed from
+the ground up with a distributed architecture (in some cases referred to as
+\textit{microservices}). Despite numerous advantages, distributed
+application design requires addressing latency and partitioning concerns that
+monolithic applications do not have (or in the case of partitioning, are not
+solvable).
+
+Our \textit{append-only datastore} addresses latency and partitioning concerns
+for append-only workloads. By explicitly distinguishing between a
+\textit{master} node and \textit{follower} nodes, we gain several advantages
+over other eventually consistent systems:
+
+\begin{enumerate}
+	\item Clients communicate only with their nearest follower for extremely
+		low latency.
+	\item Followers provide linearizability for client appends on a single
+		key. That is, all clients writing to the same key of the same
+		follower will see linearized writes.
+	\item Eventual consistency is minimally disruptive to clients. A client
+		may read the data for key $k$ and receive data consisting of 2
+		writes: $w_1, w_2$. Write $w_x$ sent through another follower
+		preserves the ordering of $w_1$ and $w_2$, so the system
+		eventually returns $w_x, w_1, w_2$, $w_1, w_x, w_2$, or $w_1,
+		w_2, w_x$ when read.  Because of (2), writes directly to the
+		same follower are linearizable.
+\end{enumerate}
+
+We also gain the advantages of some more traditionally eventually consistent
+systems, such as great partition tolerance \cite{dynamo}. Together, these
+properties are highly desirable for many append-only applications. Consider the
+following:
+
+\begin{itemize}
+	\item \textbf{Chat} - Chat applications are inherently append-only. Users
+		in different parts of the globe expect near-instantaneous (only
+		a few seconds) latency. Users in the same building, however,
+		expect to receive messages instantaneously. Think of a global
+		company: teammates send messages and links to nearby teammates,
+		expecting them to arrive immediately. Consider a chatroom
+		occupied by a team in North American and a team in Asia. Each
+		team member sees their local peers' messages immediately and in
+		exactly the order that they are sent. It's acceptable for the a
+		message from the other team to take a bit longer and appear
+		slightly out of order in the logged chat.
+	\item \textbf{Distributed logging} - Consider a distributed application
+		running in multiple data centers on multiple continents. The
+		application is monitored in real time in each datacenter. Logs
+		are write-only, and critical for monitoring. By using the
+		append-only datastore for logging, the monitoring service:
+		\begin{itemize}
+			\item Continues operating when a datacenter is
+				partitioned from the rest of the system.
+			\item Receives logs with extremely low latency from the
+				nearby monitored application.
+			\item Has an eventually consistent log that can be
+				stored for later analysis and debugging.
+		\end{itemize}
+	% \item \textbf{Distributed data processing}
+\end{itemize}
 
 \section{Design}
-We made it go super fast like vroom-vroom.
+We made it go super fast.
 
 \subsection{Leader}
-It synchronizes n'stuff.
+% Reference GFS and not transporting the data itself.
+It synchronizes stuff.
 
@@
 \section{Related Work}
+% Talk about Kafka
 This is the most filler of all filler sections. Look, a reference! \cite{dynamo}

### 2020-05-31 follower/follower.go
Make sync request asynchronously (#25)

@@ -2,7 +2,6 @@ package main
 
 import (
 	"context"
-	"errors"
 	"flag"
 	"fmt"
 	"log"
@@ -192,7 +191,7 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 func (f *follower) Sync(ctx context.Context, req *pb.SyncRequest) (*pb.SyncResponse, error) {
 	log.Printf("Received Sync request %v", req)
 	if _, ok := f.store[req.Key]; !ok {
-		return nil, errors.New(fmt.Sprint("sync message discarded because the key %v is not found", req.Key))
+		return nil, fmt.Errorf("sync message discarded because the key %v is not found", req.Key)
 	}
 
 	askForVers := req.AskFor.Versions
@@ -355,7 +354,9 @@ func (f *follower) handleUpdate(key string) error {
 					backupID:    updateResp.PreBackup.FollowerId,
 					backupAddr:  updateResp.PreBackup.Address,
 				}
-				data.syncReqChan <- s
+				go func() {
+					data.syncReqChan <- s
+				}()
 				log.Printf("sync request is in the channel")
 			}
 			data.buffer = data.buffer[:0]
@@ -395,7 +396,9 @@ func (f *follower) Notify(ctx context.Context, req *pb.NotifyRequest) (*pb.Notif
 		backupAddr:  req.Backup.Address,
 	}
 	log.Printf("sync request is in the channel, %v", s)
-	data.syncReqChan <- s
+	go func() {
+		data.syncReqChan <- s
+	}()
 
 	// TODO: Need to think about if this is safe.
 	// The version is updated after handling the braodcasting notification.

### 2020-06-01 paper/paper.tex
Did a bunch of the follower

@@
 Our \textit{append-only datastore} addresses latency and partitioning concerns
-for append-only workloads. By explicitly distinguishing between a
-\textit{master} node and \textit{follower} nodes, we gain several advantages
+for append-only workloads. It is designed to run as a service distributed
+globally across multiple data centers. By explicitly distinguishing between a
+\textit{leader} node and \textit{follower} nodes, we gain several advantages
 over other eventually consistent systems:
@@
 \section{Design}
-We made it go super fast.
+% TODO: Add architecture diagram.
+Nodes in the append-only datastore are classified in a simple hierarchy as seen
+in TODO(add the architecture diagram). A single \textit{leader} directs multiple
+\textit{followers}. Clients connnect to and communicate with a nearby follower,
+likely running in the same datacenter, to minimize latency. Our design tolerates
+arbitary partitions between nodes and ensures an eventually consistent view of
+data.
+
+Clients are exposed the following API by followers:
+\begin{itemize}
+	\item \texttt{append(key, data)} - Appends \texttt{data} to the existing
+		data for \texttt{key}. Guarantees that \texttt{data} appears
+		after any data that the client already read for \texttt{key}.
+	\item \texttt{get(key)} - Gets the data for \texttt{key}, expressed as a
+		list of indexed values. For example, the returned data will be
+		the result of writes 1, 3, and 5, expressed as $\{w_1, w_3,
+		w_5\}$. Subsequent \texttt{get} requests will always return
+		$w_1$, $w_3$, and $w_5$ in the same relative order, but other
+		writes may be prepended, appended, or interspersed.
+\end{itemize}
+
+\subsection{Follower}
+% Explain what happens when this is partitioned.
+% Explain read-your-own-writes consistency.
+% Be sure to discuss any synchronization that happens here, as the follower is
+% responsible for synchronizing its the writes to it.
+The follower is our most complex component. Its responsibilities include:
+
+\begin{itemize}
+	\item Handing \texttt{append} and \texttt{get} requests from clients.
+	\item Ordering appends.
+	\item Updating the master about appended data.
+	\item Receiving the eventual, global ordering of data from the master
+		and imposing it on data.
+	\item Syncing with other followers to communicate updates and orderings.
+\end{itemize}
+
+% TODO: We don't actually do the update-on-get thing yet.
+Each follower stores the entire datastore locally in memory. When clients issue
+\texttt{get(k)} requests, these are served from the in-memory datastore to minimize
+response latency. It also causes the follower to ask the leader for the most
+up-to-date data for \texttt{k}, ensuring that frequently accessed data is kept
+relatively fresh.
+
+\texttt{append} requests are also fast, limited by the speed of a local disk
+write. When a client appends data to key \texttt{k}, the follower writes that
+data - along with its index - to local disk before returning to the client. This
+ensures that data is not lost upon crashing and rebooting. For increased fault
+tolerance, the append should be written to multiple disks locally or to a remote
+disk.
+
+As soon as the append is written to disk, the follower (concurrently with
+replying to the client) sends an \texttt{update} request to the leader
+containing the key being appended to and the index that the follower has
+assigned the appended data. The leader resolves and orders the append with any
+other concurrent appends from different followers (see next section), and
+responds.
+
+There are two types of responses. The first - and likely most common - case
+is a simple \texttt{SUCCESS} response. This indicates that the append committed
+successfully and the follower has up-to-date data for that key. The second is a
+\texttt{NEEDS\_UPDATE} response. This indicates that the data was appended, but
+another \texttt{update} preempted it. In this scenario, the follower changes the
+locally stored index of the appended data, and sends a \texttt{sync} request
+another follower indicated by the leader's response.
+
+Followers use the \texttt{sync} request to get data they are missing,
+effectively ``patching the holes" in their local datastore. This ensures the
+eventually consistent property of the system.
+
+In the common case, where updates do not overlap...
+
+When concurrent updates do occur...
+
+Both \texttt{update} and \texttt{sync} requests are retried in case of network
+failure. If a follower reboots during a \texttt{update}, it can inspect the disk
+at boot and see that it has appended data locally that has not been confirmed by
+the leader. If a follower reboots during a \texttt{sync}, it will have stale
+data until a \texttt{get} or \texttt{append} causes it to update. This keeps the
+system simple while retaining its consistency properties.
 
 \subsection{Leader}
+% Explain what happens when this is partitioned.
 % Reference GFS and not transporting the data itself.
 It synchronizes stuff.
-
-\subsection{Follower}
-% Be sure to discuss any synchronization that happens here, as the follower is
-% responsible for synchronizing its the writes to it.
-It also sorta synchronizes, but mostly it follows.
 
@@
 We have very distinguished clientele.
+
+\subsection{Common Case}
+
+\subsection{Corner Cases}
+% Multiple simultaneous writes from c1 and c2 to f1. How do we ensure that the
+% GL does not order w2 before w1?
 
@@
 \section{Future Work}
+% Use a raft group instead of a single node for each follower and GL.
+% Client fallback to other followers.
 Make it actually work.

### 2020-06-01 paper/paper.tex
Finished first draft of the follower

@@
 		may read the data for key $k$ and receive data consisting of 2
-		writes: $w_1, w_2$. Write $w_x$ sent through another follower
-		preserves the ordering of $w_1$ and $w_2$, so the system
-		eventually returns $w_x, w_1, w_2$, $w_1, w_x, w_2$, or $w_1,
-		w_2, w_x$ when read.  Because of (2), writes directly to the
+		writes' data: $d_1, d_2$. Write $x$ with data $d_x$ sent through another follower
+		preserves the ordering of $d_1$ and $d_2$, so the system
+		eventually returns $d_x, d_1, d_2$, $d_1, d_x, d_2$, or $d_1,
+		d_2, d_x$ when read.  Because of (2), writes directly to the
 		same follower are linearizable.
@@
 		list of indexed values. For example, the returned data will be
-		the result of writes 1, 3, and 5, expressed as $\{w_1, w_3,
-		w_5\}$. Subsequent \texttt{get} requests will always return
-		$w_1$, $w_3$, and $w_5$ in the same relative order, but other
+		the result of writes 1, 3, and 5, expressed as $\{d_1, d_3,
+		d_5\}$. Subsequent \texttt{get} requests will always return
+		$d_1$, $d_3$, and $d_5$ in the same relative order, but other
 		writes may be prepended, appended, or interspersed.
@@
 tolerance, the append should be written to multiple disks locally or to a remote
-disk.
+disk. Note that throughout this section when we refer to updating local storage,
+the in-memory store is also updated immediately after data is written to disk.
 
@@
 There are two types of responses. The first - and likely most common - case
-is a simple \texttt{SUCCESS} response. This indicates that the append committed
+is a simple \texttt{SUCCESS} response. This indicates that the append completed
 successfully and the follower has up-to-date data for that key. The second is a
@@
 
-In the common case, where updates do not overlap...
+\subsubsection{Common Case}
+When updates do not overlap, the follower handles \texttt{append} requests as
+follows:
 
-When concurrent updates do occur...
+\begin{enumerate}
+	\item Write the append to local storage along with the index \texttt{i}
+		the follower believes applies to the append.
+	\item Respond to the client that the request was successful.
+	\item Concurrently with (2), send an \texttt{update} to the leader that
+		an append occured on key \texttt{k} at index \texttt{i}.
+	\item The leader responds with \texttt{SUCCESS}.
+	\item The follower updates local disk to indicate that the append has
+		been commited at index \texttt{i} by the leader.
+\end{enumerate}
 
+The data for key \texttt{k} has simply benn appended to, changing from $\{d_a,
+d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ..., d_{i-1}, d_i\}$, where $a < b <
+c < i$. Note that there may be data ``holes" between $a$, $b$, $c$, and $i$ not
+yet filled in by either a \texttt{get} or \texttt{sync}.
+
+\subsubsection{Conccurent Updates}
+When concurrent updates occur:
+
+% TODO: There's some ugly mixing of texttt and math notation here, particularly
+% around i'.
+\begin{enumerate}
+	\item The same as steps 1-3 above.
+	\item The leader responds with \texttt{NEEDS\_UPDATE}.
+	\item The follower updates local disk to indicate that the append has
+		been commited at index \texttt{i`} by the leader.
+	\item The follower sends a \texttt{sync} request to the follower that
+		appended the data at index \texttt{i`-1}, as indicated in the
+		leader's response.
+	\item The follower receives and stores on local disk the data it was
+		missing, including what the leader deemed to be the ``winner" of
+		index \texttt{i}.
+\end{enumerate}
+
+The data for key \texttt{k} in this case is an append and a reordering,
+changing from $\{d_a, d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ...,
+d_{i-1}, d_i, d_{i'}\}$, where $a < b < c < i$ and as above may have more holes
+to fill.
+
+% TODO: We don't yet implement the early return of data.
+It is possible for a chain of \texttt{sync} requests to form when more than 2
+followers append a key at the same time. If follower $a$ starts syncing with
+follower $b$, which is already waiting on follower $c$ to respond to a
+\texttt{sync}, $b$ will send what is knows to $a$. $a$ then retries to get
+missing data, at which point $b$ waits to respond until it has new data to send.
+
+\subsubsection{Follower Fault Tolerance}
 Both \texttt{update} and \texttt{sync} requests are retried in case of network
@@
 
-\subsection{Common Case}
-
 \subsection{Corner Cases}
@@
 % GL does not order w2 before w1?
+% What if the sync request can be partially satisfied. Do we not respond, or
+% respond partially and expect another sync request from the same follower?
 

### 2020-06-01 paper/paper.tex
Finished the rough draft.

@@
 We also gain the advantages of some more traditionally eventually consistent
-systems, such as great partition tolerance \cite{dynamo}. Together, these
+systems, such as great partition tolerance \cite{Dynamo}. Together, these
 properties are highly desirable for many append-only applications. Consider the
@@
 	\item Ordering appends.
-	\item Updating the master about appended data.
-	\item Receiving the eventual, global ordering of data from the master
+	\item Updating the leader about appended data.
+	\item Receiving the eventual, global ordering of data from the leader
 		and imposing it on data.
@@
 tolerance, the append should be written to multiple disks locally or to a remote
-disk. Note that throughout this section when we refer to updating local storage,
-the in-memory store is also updated immediately after data is written to disk.
+disk as well. Note that throughout this section when we refer to updating local
+storage, the in-memory store is also updated immediately after data is written
+to disk.
 
@@
 effectively ``patching the holes" in their local datastore. This ensures the
-eventually consistent property of the system.
+eventually consistent property of the system. \texttt{sync} requests also signal
+what data the sender has to the recipient, enabling the recipeint to also patch
+holes in its local datastore.
 
@@
 \subsection{Leader}
-% Explain what happens when this is partitioned.
-% Reference GFS and not transporting the data itself.
-It synchronizes stuff.
+% TODO: Explain what happens when this is partitioned. Look at gfs paper.
+The leader is responsible for globally ordering writes for each key. Like GFS's
+master \cite{GFS}, it only stores metadata rather than the datastore itself.
+This decreases network utilization, as data itself is never sent.  Also, the
+burden on disks is lighter per update, increasing writes per second and the
+average lifetime of each disk. Like GFS, the simplicity of a single master
+simplifies our implementation and the rest of our design, both of which increase
+the overall stability of the system.
+
+% TODO: Add the nonce.
+The leader answers \texttt{update} requests from followers. These requests
+contain the key being appended to, the index at which the follower is trying
+to append, and a nonce to prevent erroneous double-appends. The leader
+authoritatively orders appends by keeping a map of key to current index. If a
+mapping exists from key $k$ to index $i$, an update specifying a write to index
+$i+1$ increments $i$ and returns \texttt{SUCCESS}. The leader writes all map
+updates to disk before responding.
+
+An \texttt{update} to $i' < i$ indicates that the sender is not synced to index
+$i$. In this case the leader increments its index and reassigns the append the
+index of $i+1$. It communicates this back to the follower, and tells it to
+sync with the follower that originated the append at index $i$.
+
+Along with the current index, the leader's map also contains the follower that
+originated the append at that index. This allows followers to be directed at the
+freshest information whether
+
+\subsubsection{Broadcasting Updates}
+Although the system as described is eventually consistent, it may lead to
+unacceptably stale data. An updated value might get synced between followers
+$a$ and $b$, but follower $c$ would miss the update entirely. If a value is
+updated and 3 days later follower $c$ syncs via a \texttt{get} or
+\texttt{append}, $c$'s client will get 3-day-old data.
+
+To prevent excessive data staleness, the leader is able to broadcast updates to
+every follower. Depending on the workload, broadcasting every update may slow
+down the system and excessively consume bandwidth. Thus users can pass a flag
+specifying whether they want broadcasts for every update or periodic broadcasts.
+New broadcast behavior is easy to add, as it requires implementing a
+single-method Go interface that can make use of the same utility functions we
+used to write our broadcasters.
+
+\subsubsection{Leader Fault Tolerance}
+Because the leader writes all updates to disk, it tolerates crashes and reboots.
+For greater fault tolerance, it should write to multiple disks or a remote disk
+as well.
+
+The system continues to function when the leader is partitioned, but followers
+and thus clients do not receive updates from other followers. As long as the
+partition is eventually healed, the system will propgate information correctly
+and self-heal. If for some reason there is a permanent partition, it must be
+manually worked around by changing the cluster configuration (i.e. the set of
+nodes in the system).
 
 \subsection{Clients}
-% Describe the API and guarantees exposed to clients.
-We have very distinguished clientele.
+The append-only datastore presents clients with a simple interface consisting of
+the \texttt{get} and \texttt{append} API described above. The consistency model
+guarantees that clients:
 
-\subsection{Corner Cases}
-% Multiple simultaneous writes from c1 and c2 to f1. How do we ensure that the
-% GL does not order w2 before w1?
-% What if the sync request can be partially satisfied. Do we not respond, or
-% respond partially and expect another sync request from the same follower?
+\begin{itemize}
+	\item Can read their own appends immiedately.
+	\item Never see data reordered relatively.
+	\item Communicate only with a fast, nearby follower, and no other nodes
+		in the system.
+	\item Can always read and write, even when their follower is partitioned
+		from the rest of the system.
+\end{itemize}
 
-\section{Implementation}
+The sacrifices made are:
+
+\begin{itemize}
+	\item Consistency is eventual, not stronger.
+	\item We do not support random writes.
+	\item The system continues to work but may provide stale data in the
+		event of a leader failure.
+\end{itemize}
 
 \section{Evaluation}
-We tested using some super cool gcloud stuff.
+The system is written entirely in Go and can be run manually or distributed and
+run with Docker. Nodes communicate via GRPC.
+
+We ran tests with a cluster of Google Compute Engine virtual
+machines. The cluster consisted of:
+
+\begin{itemize}
+	\item A leader in Seoul, Korea.
+	\item A follower and client in Frankfurt, Germany.
+	\item A follower and client in Los Angeles, USA.
+	\item A follower and client in São Paulo, Brazil.
+\end{itemize}
 
 \subsection{Workload 1}
-Oh wow, we're so fast.
+We should probably read and write a lot.
 
 \subsection{Workload 2}
-Oh no, we're unexpectedly not that fast.
+Let's see what happens if we kill a follower.
 
 \section{Related Work}
-% Talk about Kafka
-This is the most filler of all filler sections. Look, a reference! \cite{dynamo}
+Amazon's Dynamo \cite{Dynamo} supports always-writable semantics, high partition
+tolerance, and laser-focuses on low-latency operation. Unlike our system, Dynamo
+is a key-value store. It also exposes a great deal of complexity to developers,
+who have to tailor their use of Dynamo such that conflicts are resolvable and
+must manually implement conflict resolution in clients. Our system does not
+allow for conflicts.
+
+Google's GFS \cite{GFS} is also optimized for append-heavy workloads and uses a
+single master for simplicity and intelligent coordination. It supports random
+writes as well. However, it is explicitly optimized for non-latency-sensitive
+applications, and a single read can require multiple hops (the GFS master and
+chunkserver).
+
+% TODO: Make this better.
+LinkedIn's Kafka \cite{Kafka} also provides eventual delivery of large
+quantities of data (specifically logs), but contains a publish/subscription
+mechanism.
 
 \section{Future Work}
-% Use a raft group instead of a single node for each follower and GL.
-% Client fallback to other followers.
-Make it actually work.
+We could greatly increase throughput by providing a client library (rather than
+a raw GRPC interface) that is aware of the version or latest verions it holds.
+This would enable followers to selectively return only the few missing pieces of
+data a client is missing rather than all the data.
+
+To increase fault tolerance, we have discussed making each follower and the
+leader their own small cluster of concensus nodes. This would greatly increase
+fault tolerance and, while it may complicate the system, would not complicate
+clients or the protocols via which nodes communicate.
+
+Lastly, there may be cases where followers fail or are partitioned from their
+nearby clients. In these cases, we would like to explore whether clients can
+fall back to another follower.
 
 \section{Conclusions}
-I'd just like everyone to pat themselves on the back here.
+We built a datastore to provide low latency and high availability for append-only
+data storage. While this work is tailed to a specific set of workloads, there
+are numerous applications that can benefit from this approach.
+
+Our system can be used to support distributed services such as distributed
+monitoring and chat applications with high performance. Importantly, it exposes
+a simple interface allowing for simple clients. We believe that this makes it a
+useful tool in building distributed and microservice systems.
 

### 2020-06-01 paper/paper.tex

@@
 in TODO(add the architecture diagram). A single \textit{leader} directs multiple
-\textit{followers}. Clients connnect to and communicate with a nearby follower,
+\textit{followers}. Clients connect to and communicate with a nearby follower,
 likely running in the same datacenter, to minimize latency. Our design tolerates
-arbitary partitions between nodes and ensures an eventually consistent view of
+arbitrary partitions between nodes and ensures an eventually consistent view of
 data.
@@
 eventually consistent property of the system. \texttt{sync} requests also signal
-what data the sender has to the recipient, enabling the recipeint to also patch
+what data the sender has to the recipient, enabling the recipient to also patch
 holes in its local datastore.
@@
 	\item Concurrently with (2), send an \texttt{update} to the leader that
-		an append occured on key \texttt{k} at index \texttt{i}.
+		an append occurred on key \texttt{k} at index \texttt{i}.
 	\item The leader responds with \texttt{SUCCESS}.
 	\item The follower updates local disk to indicate that the append has
-		been commited at index \texttt{i} by the leader.
+		been committed at index \texttt{i} by the leader.
 \end{enumerate}
 
-The data for key \texttt{k} has simply benn appended to, changing from $\{d_a,
+The data for key \texttt{k} has simply been appended to, changing from $\{d_a,
 d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ..., d_{i-1}, d_i\}$, where $a < b <
@@
 
-\subsubsection{Conccurent Updates}
+\subsubsection{Concurrent Updates}
 When concurrent updates occur:
@@
 	\item The follower updates local disk to indicate that the append has
-		been commited at index \texttt{i`} by the leader.
+		been committed at index \texttt{i`} by the leader.
 	\item The follower sends a \texttt{sync} request to the follower that
@@
 and thus clients do not receive updates from other followers. As long as the
-partition is eventually healed, the system will propgate information correctly
+partition is eventually healed, the system will propagate information correctly
 and self-heal. If for some reason there is a permanent partition, it must be
@@
 \begin{itemize}
-	\item Can read their own appends immiedately.
+	\item Can read their own appends immediately.
 	\item Never see data reordered relatively.
@@
 We could greatly increase throughput by providing a client library (rather than
-a raw GRPC interface) that is aware of the version or latest verions it holds.
+a raw GRPC interface) that is aware of the indeces or latest index it holds.
 This would enable followers to selectively return only the few missing pieces of
@@
 To increase fault tolerance, we have discussed making each follower and the
-leader their own small cluster of concensus nodes. This would greatly increase
+leader their own small cluster of consensus nodes. This would greatly increase
 fault tolerance and, while it may complicate the system, would not complicate

### 2020-06-01 follower/follower.go
Also create new entry upon receiving a new key from the broadcasted Notify request (#27)

@@ -123,24 +123,28 @@ func (f *follower) stop() {
 	f.conn.Close()
 }
 
+func (f *follower) initNewKey(key string) {
+	// A new key is received. Initialize a new entry in the store.
+	f.store[key] = &data{
+		updateReqChan:   make(chan *lpb.UpdateRequest, 100),
+		syncReqChan:     make(chan *syncRequest, 100),
+		syncReqQueue:    make([]*syncRequest, 0),
+		done:            make(chan bool),
+		buffer:          make([]string, 0),
+		versionToValues: make(map[int64][]string),
+	}
+	// Starts channels
+	go f.handleUpdate(key)
+	go f.handleSync(key)
+}
+
 func (f *follower) Put(ctx context.Context, req *pb.PutRequest) (*pb.PutResponse, error) {
 	// Storing the data
 	log.Printf("Received Put request %v", req)
 
 	// Get the current version (or 0 if it's the first).
 	if _, ok := f.store[req.Key]; !ok {
-		// A new key is received. Initialize a store entry.
-		f.store[req.Key] = &data{
-			updateReqChan:   make(chan *lpb.UpdateRequest, 100),
-			syncReqChan:     make(chan *syncRequest, 100),
-			syncReqQueue:    make([]*syncRequest, 0),
-			done:            make(chan bool),
-			buffer:          make([]string, 0),
-			versionToValues: make(map[int64][]string),
-		}
-		// Starts channels
-		go f.handleUpdate(req.Key)
-		go f.handleSync(req.Key)
+		f.initNewKey(req.Key)
 	}
 
 	// newData := &value{val: req.Value, version: v}
@@ -370,7 +374,8 @@ func (f *follower) handleUpdate(key string) error {
 func (f *follower) Notify(ctx context.Context, req *pb.NotifyRequest) (*pb.NotifyResponse, error) {
 	log.Printf("Got a notification, %v", req)
 	if _, ok := f.store[req.Key]; !ok {
-		return nil, fmt.Errorf("key not found %v", req.Key)
+		// Got a new Key
+		f.initNewKey(req.Key)
 	}
 
 	data := f.store[req.Key]

### 2020-06-02 follower/follower.go
allow binding to all 0.0.0.0 for setting up a non-localhost env.

@@ -426,7 +426,7 @@ func main() {
 	}
 
 	// Create and start the follower service.
-	lis, err := net.Listen("tcp", util.FormatServiceAddress(followerAddress))
+	lis, err := net.Listen("tcp", util.FormatBindAddress(followerAddress))
 	if err != nil {
 		log.Fatalf("failed to listen: %v", err)
 	}

### 2020-06-05 client/client.go
Perf test continue

@@ -4,9 +4,11 @@ import (
 	"context"
 	"flag"
 	"fmt"
+	"io"
 	"io/ioutil"
 	"log"
 	"math/rand"
+	"os"
 	"sort"
 	"sync"
 	"time"
@@ -23,6 +25,7 @@ import (
 const followerIDFlag = "follower_id"
 const timeoutFlag = "timeout"
 const modeFlag = "mode"
+const logOutputFlag = "logOutput"
 
 var (
 	followerID = flag.String(followerIDFlag, "", "The Follower to talk to")
@@ -30,12 +33,15 @@ var (
 	timeout    = flag.Int(timeoutFlag, 5, "Timeout, in seconds, when connecting to the follower")
 	testTime   = flag.Int("testtime", 30, "How long to send updates for, in seconds")
 
-	threadCount  = flag.Int("threadcount", 1, "The number of threads that sends a number of requests")
+	threadCount  = flag.Int("threadcount", 2, "The number of threads that sends a number of requests")
 	requestCount = flag.Int("requestcount", 10, "The number of requests to send")
 	testKeys     = []string{"key-0", "key-1", "key-2", "key-3", "key-4", "key-5", "key-6", "key-7", "key-8", "key-9"}
 	// testKeys   = []string{"key-0"}
 	valueCount = 0
 	countLock  = sync.Mutex{}
+
+	logOutput = flag.String(logOutputFlag, "log-only", "Path to the log file")
+	logger    *log.Logger
 )
 
 // sentRequest
@@ -50,12 +56,12 @@ func sendRequest(ctx context.Context, c flpb.FollowerClient, t time.Time) {
 	valueCount++
 	countLock.Unlock()
 
-	log.Printf("sending request %v", req)
+	logger.Printf("sending request %v", req)
 	resp, err := c.Put(ctx, req)
 	if err != nil {
-		log.Fatalf("%v.Put(_) = _, %v: ", c, err)
+		logger.Fatalf("%v.Put(_) = _, %v: ", c, err)
 	}
-	log.Println(resp)
+	logger.Println(resp)
 }
 
 // sentRequest
@@ -67,17 +73,17 @@ func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time) 
 		Value: fmt.Sprintf("client-%v-%v-value-%s", *followerID, k, uuid.New()),
 	}
 
-	log.Printf("sending request %v", req)
[…]

### 2020-06-05 follower/follower.go

@@ -30,6 +30,7 @@ const logOutputFlag = "logOutput"
 var (
 	followerID            = flag.String(followerIDFlag, "", "The Follower's Unique ID")
 	timeout               = flag.Int(timeoutFlag, 5, "Timeout, in seconds, when connecting to the leader")
+	connMapMutex          = sync.Mutex{}
 	followerConnectionMap = make(map[string]pb.FollowerClient)
 	logOutput             = flag.String(logOutputFlag, "log-only", "Path to the log file")
 	logger                *log.Logger
@@ -148,7 +149,6 @@ func (f *follower) initNewKey(key string) {
 
 func (f *follower) Put(ctx context.Context, req *pb.PutRequest) (*pb.PutResponse, error) {
 	// Storing the data
-	logger.Printf("Received Put request %v", req)
 	f.storeMutex.Lock()
 	// Get the current version (or 0 if it's the first).
 	if _, ok := f.store[req.Key]; !ok {
@@ -162,8 +162,8 @@ func (f *follower) Put(ctx context.Context, req *pb.PutRequest) (*pb.PutResponse
 
 	go f.sendUpdate(req.Key)
 
+	logger.Printf("Received Put request %v, key %v is on version %v now", req, req.Key, f.store[req.Key].latestVersion)
 	f.storeMutex.Unlock()
-	logger.Println(res)
 
 	// TODO: write the data to the local disk
 	// Infomation that needs to be written:
@@ -183,15 +183,15 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 	var result []string
 	// First get the data from the store, which are all of the synced data.
 	if d, ok := f.store[req.Key]; ok {
-		var keys []int64
-		for k := range d.versionToValues {
-			keys = append(keys, k)
+		var versions []int64
+		for v := range d.versionToValues {
+			versions = append(versions, v)
 		}
 
 		sort.Slice(keys, func(i, j int) bool { return keys[i] < keys[j] })
 
 		var v []string
-		for _, k := range keys {
+		for _, k := range versions {
 			v = append(v, d.versionToValues[k]...)
 		}
 		result = append(result, v...)
@@ -212,6 +212,7 @@ func (f *follower) Sync(ctx context.Context, req *pb.SyncRequest) (*pb.SyncRespo
 	askForVers := req.AskFor.Versions
 	incomingData := req.MyData
[…]

### 2020-06-05 leader/leader.go

@@ -150,15 +150,13 @@ func (l *leader) Update(ctx context.Context, req *pb.UpdateRequest) (*pb.UpdateR
 
 	// Update the follower info. Swtich the current primary to the input follower info.
 	prePrimary := *l.keyVersionMap.data[req.Key].fInfo
-	logger.Printf("preprimary is %v, id %v", prePrimary.followerAddr.Port, prePrimary.followerID)
 
 	l.keyVersionMap.data[req.Key].fInfo.followerAddr = req.FollowerAddress
 	l.keyVersionMap.data[req.Key].fInfo.followerID = req.FollowerId
-	logger.Printf("new preprimary is %v id %v", l.keyVersionMap.data[req.Key].fInfo.followerAddr.Port, l.keyVersionMap.data[req.Key].fInfo.followerID)
-	logger.Printf("preprimary again is %v, id %v", prePrimary.followerAddr.Port, prePrimary.followerID)
 
 	// Set the result
 	result := pb.UpdateResult_SUCCESS
+	logger.Printf("key %v, leader's version is %d, input follower's version is %d", req.Key, l.keyVersionMap.data[req.Key].version, req.Version)
 	if l.keyVersionMap.data[req.Key].version > req.Version && prePrimary.followerID != req.FollowerId {
 		result = pb.UpdateResult_NEED_SYNC
 	}
@@ -249,7 +247,7 @@ func main() {
 		logger = log.New(os.Stdout, "", log.LstdFlags)
 	case "both":
 		logFile, err := os.OpenFile("leader.log",
-			os.O_APPEND|os.O_CREATE|os.O_WRONLY, 0644)
+			os.O_RDWR|os.O_CREATE|os.O_TRUNC, 0644)
 		if err != nil {
 			log.Println(err)
 		}
@@ -262,7 +260,7 @@ func main() {
 	default:
 		// Write to log file only
 		logFile, err := os.OpenFile("leader.log",
-			os.O_APPEND|os.O_CREATE|os.O_WRONLY, 0644)
+			os.O_RDWR|os.O_CREATE|os.O_TRUNC, 0644)
 		if err != nil {
 			log.Println(err)
 		}

### 2020-06-05 test/perf_test.sh
Perf test continue (#30)

@@ -1,5 +1,9 @@
 #!/bin/bash
 
+THREADCOUNT=1
+REQUESTSCOUNT=400
+TESTTIME=50
+
 function startFollowers {
   # TODO: use the docker container to bring up the env
   config_file=${1}
@@ -32,18 +36,19 @@ function startFollowers {
 
 function startClients {
   config_file=${1}
+  # Perf test parameters
 
-  bin/client --follower_id=0 --mode=perf --config_file=${config_file}  &
+  bin/client --follower_id=0 --mode=perf --config_file=${config_file} --threadcount=$THREADCOUNT --requestcount=$REQUESTSCOUNT --testtime=$TESTTIME &
   CLIENT0_PID=$!
-  bin/client --follower_id=1 --mode=perf --config_file=${config_file} > c-1.log &
+  bin/client --follower_id=1 --mode=perf --config_file=${config_file} --threadcount=$THREADCOUNT --requestcount=$REQUESTSCOUNT --testtime=$TESTTIME &
   CLIENT1_PID=$!
-  bin/client --follower_id=2 --mode=perf --config_file=${config_file} > c-2.log &
+  bin/client --follower_id=2 --mode=perf --config_file=${config_file} --threadcount=$THREADCOUNT --requestcount=$REQUESTSCOUNT --testtime=$TESTTIME &
   CLIENT2_PID=$!
-  bin/client --follower_id=3 --mode=perf --config_file=${config_file} > c-3.log &
+  bin/client --follower_id=3 --mode=perf --config_file=${config_file} --threadcount=$THREADCOUNT --requestcount=$REQUESTSCOUNT --testtime=$TESTTIME &
   CLIENT3_PID=$!
-  bin/client --follower_id=4 --mode=perf --config_file=${config_file} > c-4.log &
+  bin/client --follower_id=4 --mode=perf --config_file=${config_file} --threadcount=$THREADCOUNT --requestcount=$REQUESTSCOUNT --testtime=$TESTTIME &
   CLIENT4_PID=$!
-  bin/client --follower_id=5 --mode=perf --config_file=${config_file} > c-5.log &
+  bin/client --follower_id=5 --mode=perf --config_file=${config_file} --threadcount=$THREADCOUNT --requestcount=$REQUESTSCOUNT --testtime=$TESTTIME &
   CLIENT5_PID=$!
 
   echo "Clients PIDs"
@@ -70,9 +75,11 @@ echo "Followers started"
 
[…]

### 2020-06-05 client/client.go
Add time measurement

@@ -40,6 +40,9 @@ var (
 	valueCount = 0
 	countLock  = sync.Mutex{}
 
+	elapsedLock = sync.Mutex{}
+	sendElapsed = int64(0)
+
 	logOutput = flag.String(logOutputFlag, "log-only", "Path to the log file")
 	logger    *log.Logger
 )
@@ -74,7 +77,12 @@ func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time) 
 	}
 
 	logger.Printf("sending request %v", req)
+	start := time.Now()
 	_, err := c.Put(ctx, req)
+	elapsed := time.Since(start)
+	elapsedLock.Lock()
+	sendElapsed += elapsed.Microseconds()
+	elapsedLock.Unlock()
 	if err != nil {
 		logger.Fatalf("%v.Put(_) = _, %v: ", c, err)
 	}
@@ -105,10 +113,13 @@ func getData(c flpb.FollowerClient) string {
 	logger.Printf("********************************")
 	var count int
 	var result string
+	start := time.Now()
+
 	for _, k := range testKeys {
 		req := &flpb.GetRequest{
 			Key: k,
 		}
+
 		resp, err := c.Get(ctx, req)
 		if err != nil {
 			logger.Fatalf("%v.Get(_) = _, %v: ", c, err)
@@ -118,7 +129,9 @@ func getData(c flpb.FollowerClient) string {
 		r := fmt.Sprintf("%v: [%v]\n\n\n", k, resp.Values)
 		result += r
 	}
-	logger.Printf("get %d values in total", count)
+	elapsed := time.Since(start)
+	perRequest := float64(elapsed.Microseconds()) / float64(len(testKeys))
+	logger.Printf("get %d values in total, elapsed %v, per request %v microseconds", count, elapsed, perRequest)
 	logger.Printf("********************************")
 	return result
 }
@@ -144,6 +157,7 @@ func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int)
 	ticker := time.NewTicker(50 * time.Millisecond)
 	defer ticker.Stop()
 	done := make(chan bool)
+	count := *requestCount
 	for {
 		select {
 		case <-done:
@@ -151,9 +165,8 @@ func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int)
 			return
 		case <-ticker.C:
 			go sendConcurRequest(ctx, c, time.Now())
[…]

### 2020-06-05 follower/follower.go
Add time measurement (#31)

@@ -188,7 +188,7 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 			versions = append(versions, v)
 		}
 
-		sort.Slice(keys, func(i, j int) bool { return keys[i] < keys[j] })
+		sort.Slice(versions, func(i, j int) bool { return versions[i] < versions[j] })
 
 		var v []string
 		for _, k := range versions {

### 2020-06-06 client/client.go
Update config

@@ -68,12 +68,12 @@ func sendRequest(ctx context.Context, c flpb.FollowerClient, t time.Time) {
 }
 
 // sentRequest
-func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time) {
+func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time, thread int) {
 	rand.Seed(time.Now().UnixNano())
 	k := testKeys[rand.Intn(len(testKeys))]
 	req := &flpb.PutRequest{
 		Key:   k,
-		Value: fmt.Sprintf("client-%v-%v-value-%s", *followerID, k, uuid.New()),
+		Value: fmt.Sprintf("client-%v-%v-t-%d-%s", *followerID, k, thread, uuid.New()),
 	}
 
 	logger.Printf("sending request %v", req)
@@ -81,6 +81,7 @@ func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time) 
 	_, err := c.Put(ctx, req)
 	elapsed := time.Since(start)
 	elapsedLock.Lock()
+	valueCount++
 	sendElapsed += elapsed.Microseconds()
 	elapsedLock.Unlock()
 	if err != nil {
@@ -111,7 +112,7 @@ func getData(c flpb.FollowerClient) string {
 	ctx := context.Background()
 	// var result = make([][]string, 10)
 	logger.Printf("********************************")
-	var count int
+	var received int
 	var result string
 	start := time.Now()
 
@@ -124,14 +125,14 @@ func getData(c flpb.FollowerClient) string {
 		if err != nil {
 			logger.Fatalf("%v.Get(_) = _, %v: ", c, err)
 		}
-		count += len(resp.Values)
+		received += len(resp.Values)
 		sort.Strings(resp.Values)
 		r := fmt.Sprintf("%v: [%v]\n\n\n", k, resp.Values)
 		result += r
 	}
 	elapsed := time.Since(start)
 	perRequest := float64(elapsed.Microseconds()) / float64(len(testKeys))
-	logger.Printf("get %d values in total, elapsed %v, per request %v microseconds", count, elapsed, perRequest)
+	logger.Printf("get %d values in total, elapsed %v, per request %v microseconds", received, elapsed, perRequest)
 	logger.Printf("********************************")
 	return result
 }
@@ -164,7 +165,7 @@ func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int)
[…]

### 2020-06-06 test/perf_test.sh

@@ -1,8 +1,8 @@
 #!/bin/bash
 
-THREADCOUNT=1
-REQUESTSCOUNT=400
-TESTTIME=50
+THREADCOUNT=2
+REQUESTSCOUNT=10
+TESTTIME=30
 
 function startFollowers {
   # TODO: use the docker container to bring up the env

### 2020-06-06 leader/leader.go
Remove some log

@@ -212,7 +212,6 @@ func (l *leader) broadcasting() {
 			logger.Println("done broadcasting")
 			return
 		case <-ticker.C:
-			logger.Println("broadcast...")
 			for id, fc := range l.followers {
 				for k, vi := range l.keyVersionMap.data {
 					if id == vi.fInfo.followerID {

### 2020-06-06 test/perf_test.sh
Config internal ip (#35)

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 THREADCOUNT=2
-REQUESTSCOUNT=10
+REQUESTSCOUNT=200
 TESTTIME=30
 
 function startFollowers {

### 2020-06-06 client/client.go
Some fixes

@@ -73,11 +73,10 @@ func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time, 
 	k := testKeys[rand.Intn(len(testKeys))]
 	req := &flpb.PutRequest{
 		Key:   k,
-		Value: fmt.Sprintf("client-%v-%v-t-%d-%s", *followerID, k, thread, uuid.New()),
+		Value: fmt.Sprintf("c-%v-t%d-%v-%s", *followerID, thread, k, uuid.New()),
 	}
-
-	logger.Printf("sending request %v", req)
 	start := time.Now()
+	logger.Printf("sending request %v, at %v", req, start.UnixNano())
 	_, err := c.Put(ctx, req)
 	elapsed := time.Since(start)
 	elapsedLock.Lock()
@@ -185,7 +184,7 @@ func perfTest(c flpb.FollowerClient) {
 
 	perRequest := float64(sendElapsed) / float64(*requestCount)
 	logger.Printf("********************************")
-	logger.Printf("sent %d values in total, elapsed %v us, per request %v us", (*requestCount)*(*threadCount), sendElapsed, perRequest)
+	logger.Printf("sent %d values in total, elapsed %v us, per request %v us", valueCount, sendElapsed, perRequest)
 	logger.Printf("********************************")
 
 	result := getData(c)
@@ -252,6 +251,13 @@ func main() {
 		runClient(client)
 	case "perf":
 		perfTest(client)
+	case "getdata":
+		result := getData(client)
+		fileName := fmt.Sprintf("./current-data-for-%s", *followerID)
+		err := ioutil.WriteFile(fileName, []byte(result), 0644)
+		if err != nil {
+			panic(err)
+		}
 	}
 
 }

### 2020-06-06 test/perf_test.sh
Some fixes (#36)

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 THREADCOUNT=2
-REQUESTSCOUNT=200
+REQUESTSCOUNT=600
 TESTTIME=30
 
 function startFollowers {

### 2020-06-07 client/client.go
Fix some bugs

@@ -152,32 +152,42 @@ func runClient(client flpb.FollowerClient) {
 	getData(client)
 }
 
-func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int) {
+func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int, wg *sync.WaitGroup) {
 	logger.Printf("this is routine %d", t)
-	ticker := time.NewTicker(50 * time.Millisecond)
-	defer ticker.Stop()
-	done := make(chan bool)
-	count := *requestCount
-	for {
-		select {
-		case <-done:
-			logger.Println("done perf test sending requests")
-			return
-		case <-ticker.C:
-			go sendConcurRequest(ctx, c, time.Now(), t)
-			count--
-			if count == 0 {
-				done <- true
-			}
-		}
+	defer wg.Done()
+	// ticker := time.NewTicker(50 * time.Millisecond)
+	// defer ticker.Stop()
+	// done := make(chan bool)
+	// count := *requestCount
+	// // for {
+	// // 	select {
+	// // 	case <-done:
+	// // 		logger.Println("done perf test sending requests")
+	// // 		return
+	// // 	case <-ticker.C:
+	// // 		go sendConcurRequest(ctx, c, time.Now(), t)
+	// 		count--
+	// 		if count == 0 {
+	// 			done <- true
+	// 		}
+	// 	}
+	// }
+
+	for i := 0; i < *requestCount; i++ {
+		sendConcurRequest(ctx, c, time.Now(), t)
 	}
 }
 
 func perfTest(c flpb.FollowerClient) {
 	ctx := context.Background()
+	var wg sync.WaitGroup
 	for i := 0; i < *threadCount; i++ {
-		go sendRequestsConcurrently(ctx, c, i)
+		wg.Add(1)
+		go sendRequestsConcurrently(ctx, c, i, &wg)
 	}
+	logger.Println("wait for wait group")
+	wg.Wait()
+	logger.Println("wait done wait group")
 
 	// Wait for other test clients to finish
 	time.Sleep(time.Duration(*testTime) * time.Second)

### 2020-06-07 follower/follower.go
Fix some bugs (#37)

@@ -182,8 +182,10 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 	logger.Printf("Received a Get request %v", req)
 	var result []string
 	// First get the data from the store, which are all of the synced data.
+	f.storeMutex.Lock()
 	if d, ok := f.store[req.Key]; ok {
 		var versions []int64
+		d.versionToValueLock.Lock()
 		for v := range d.versionToValues {
 			versions = append(versions, v)
 		}
@@ -194,9 +196,11 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 		for _, k := range versions {
 			v = append(v, d.versionToValues[k]...)
 		}
+		d.versionToValueLock.Unlock()
 		result = append(result, v...)
 		result = append(result, f.store[req.Key].buffer...)
 	}
+	f.storeMutex.Unlock()
 
 	return &pb.GetResponse{
 		Values: result,
@@ -219,8 +223,8 @@ func (f *follower) Sync(ctx context.Context, req *pb.SyncRequest) (*pb.SyncRespo
 	// The version numbers should all exist in the store.
 	// Only reply the versions that being asked for.
 	for _, v := range askForVers {
-		if _, ok := mydata[v]; !ok {
-			err := fmt.Errorf("syncing not completed, missing version %v for key %v, request %v", v, req.Key, req)
+		if d, ok := mydata[v]; !ok || len(d) == 0 {
+			err := fmt.Errorf("syncing not completed, version exist? %v, data exist? %v, for key %v version %v", ok, len(d) == 0, req.Key, v)
 			logger.Print(err)
 			continue
 		}
@@ -327,20 +331,23 @@ func (f *follower) handleSync(key string) error {
 			newValues := syncResp.Value
 			f.storeMutex.Lock()
 			f.store[key].versionToValueLock.Lock()
-			myData := f.store[key].versionToValues
 			// Put each value list under its version.
 			askForVersionsLeft := make(map[int64]bool)
 			for _, v := range syncReqest.req.AskFor.Versions {
 				askForVersionsLeft[v] = true
 			}
 			for _, v := range newValues {
-				if d, ok := myData[v.Version]; ok {
+				if d, ok := f.store[key].versionToValues[v.Version]; ok {
 					err := fmt.Errorf("version number already exist, key %v, existing data %v, incomming data %v", key, d, v)
+					logger.Println(err)
+					continue
+				} else if len(v.Value) == 0 {
[…]

### 2020-06-07 leader/leader.go

@@ -184,7 +184,7 @@ func (l *leader) Update(ctx context.Context, req *pb.UpdateRequest) (*pb.UpdateR
 	// TODO: Comment it out temporarily for testing.
 	// l.broadcaster.enqueue(req.Key, resp)
 
-	logger.Printf("leader replying update response %v", resp)
+	logger.Printf("leader replying update for %v, response %v", req.Key, resp)
 	return resp, nil
 }
 

### 2020-06-07 test/perf_test.sh

@@ -1,8 +1,8 @@
 #!/bin/bash
 
-THREADCOUNT=2
-REQUESTSCOUNT=600
-TESTTIME=30
+THREADCOUNT=100
+REQUESTSCOUNT=100
+TESTTIME=8
 
 function startFollowers {
   # TODO: use the docker container to bring up the env

### 2020-06-07 follower/follower.go
Fix bugs and refactor log

@@ -155,12 +155,17 @@ func (f *follower) Put(ctx context.Context, req *pb.PutRequest) (*pb.PutResponse
 		f.initNewKey(req.Key)
 	}
 
+	// Only update when the buffer is empty, which means the leader has accepted
+	// the last update request and the buffer has been moved to the data store.
+	// So it can send a new update request for the incoming data.
+	if len(f.store[req.Key].buffer) == 0 {
+		go f.sendUpdate(req.Key)
+	}
+
 	// The incoming values are put in the buffer and waiting to be udpated by the leader.
 	f.store[req.Key].buffer = append(f.store[req.Key].buffer, req.Value)
 	// f.store[req.Key].values = append(f.store[req.Key].values, newData)
 	res := fmt.Sprintf("key value pair recieved %v: %v", req.Key, req.Value)
-
-	go f.sendUpdate(req.Key)
 
 	logger.Printf("Received Put request %v, key %v is on version %v now", req, req.Key, f.store[req.Key].latestVersion)
 	f.storeMutex.Unlock()
@@ -191,7 +196,7 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 		}
 
 		sort.Slice(versions, func(i, j int) bool { return versions[i] < versions[j] })
-
+		// logger.Printf("key %v versions are %v", req.Key, versions)
 		var v []string
 		for _, k := range versions {
 			v = append(v, d.versionToValues[k]...)
@@ -208,7 +213,7 @@ func (f *follower) Get(ctx context.Context, req *pb.GetRequest) (*pb.GetResponse
 }
 
 func (f *follower) Sync(ctx context.Context, req *pb.SyncRequest) (*pb.SyncResponse, error) {
-	logger.Printf("Received Sync request %v", req)
+	logger.Printf("Received Sync request for key %v, asking for versions %v", req.Key, req.AskFor.Versions)
 	if _, ok := f.store[req.Key]; !ok {
 		return nil, fmt.Errorf("sync message discarded because the key %v is not found", req.Key)
 	}
@@ -224,7 +229,7 @@ func (f *follower) Sync(ctx context.Context, req *pb.SyncRequest) (*pb.SyncRespo
 	// Only reply the versions that being asked for.
 	for _, v := range askForVers {
 		if d, ok := mydata[v]; !ok || len(d) == 0 {
[…]

### 2020-06-07 client/client.go
Fix bugs and refactor log (#38)

@@ -73,10 +73,10 @@ func sendConcurRequest(ctx context.Context, c flpb.FollowerClient, t time.Time, 
 	k := testKeys[rand.Intn(len(testKeys))]
 	req := &flpb.PutRequest{
 		Key:   k,
-		Value: fmt.Sprintf("c-%v-t%d-%v-%s", *followerID, thread, k, uuid.New()),
+		Value: fmt.Sprintf("value-client-%v-t%d-%v-%s", *followerID, thread, k, uuid.New()),
 	}
 	start := time.Now()
-	logger.Printf("sending request %v, at %v", req, start.UnixNano())
+	// logger.Printf("sending request %v, at %v", req, start.UnixNano())
 	_, err := c.Put(ctx, req)
 	elapsed := time.Since(start)
 	elapsedLock.Lock()

### 2020-06-07 test/perf_test.sh

@@ -1,8 +1,8 @@
 #!/bin/bash
 
 THREADCOUNT=100
-REQUESTSCOUNT=100
-TESTTIME=8
+REQUESTSCOUNT=600
+TESTTIME=4
 
 function startFollowers {
   # TODO: use the docker container to bring up the env
@@ -82,13 +82,13 @@ sleep $TESTTIME
 sleep 10
 
 echo "Kill Servers"
-kill $LEADER_PID
 kill $FOLLOWER0_PID
 kill $FOLLOWER1_PID
 kill $FOLLOWER2_PID
 kill $FOLLOWER3_PID
 kill $FOLLOWER4_PID
 kill $FOLLOWER5_PID
+kill $LEADER_PID
 
 echo "Comparing results..."
 echo "diff 0 to 1..."

### 2020-06-07 client/client.go
Periodic get

@@ -109,12 +109,10 @@ func sentRequests(client flpb.FollowerClient) {
 
 func getData(c flpb.FollowerClient) string {
 	ctx := context.Background()
-	// var result = make([][]string, 10)
 	logger.Printf("********************************")
 	var received int
 	var result string
 	start := time.Now()
-
 	for _, k := range testKeys {
 		req := &flpb.GetRequest{
 			Key: k,
@@ -131,7 +129,7 @@ func getData(c flpb.FollowerClient) string {
 	}
 	elapsed := time.Since(start)
 	perRequest := float64(elapsed.Microseconds()) / float64(len(testKeys))
-	logger.Printf("get %d values in total, elapsed %v, per request %v microseconds", received, elapsed, perRequest)
+	logger.Printf("get %d values in total, elapsed %v, per request %v microseconds, started at %v", received, elapsed, perRequest, start)
 	logger.Printf("********************************")
 	return result
 }
@@ -153,7 +151,6 @@ func runClient(client flpb.FollowerClient) {
 }
 
 func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int, wg *sync.WaitGroup) {
-	logger.Printf("this is routine %d", t)
 	defer wg.Done()
 	// ticker := time.NewTicker(50 * time.Millisecond)
 	// defer ticker.Stop()
@@ -178,6 +175,21 @@ func sendRequestsConcurrently(ctx context.Context, c flpb.FollowerClient, t int,
 	}
 }
 
+func periodicGetData(c flpb.FollowerClient) {
+	ticker := time.NewTicker(500 * time.Millisecond)
+	defer ticker.Stop()
+	done := time.After(time.Duration(*testTime) * time.Second)
+	for {
+		select {
+		case <-done:
+			logger.Println("done periodic get data")
+			return
+		case <-ticker.C:
+			getData(c)
+		}
+	}
+}
+
 func perfTest(c flpb.FollowerClient) {
 	ctx := context.Background()
 	var wg sync.WaitGroup
@@ -189,10 +201,12 @@ func perfTest(c flpb.FollowerClient) {
 	wg.Wait()
 	logger.Println("wait done wait group")
 
+	periodicGetData(c)
+
 	// Wait for other test clients to finish
 	time.Sleep(time.Duration(*testTime) * time.Second)
 
-	perRequest := float64(sendElapsed) / float64(*requestCount)
+	perRequest := float64(sendElapsed) / (float64(*requestCount) * float64(*threadCount))
[…]

### 2020-06-08 paper/paper.tex
Proofread
Fixes to grammar, spelling, a little meaning.

--- 2020-06-01 paper/paper.tex
@@
 \documentclass[11pt,english,twocolumn]{article}
+
+\usepackage{graphicx}
+\usepackage{subfigure}
 
@@
 \author{
-	Mingyan Zhao
+	Mingyan Zhao \\
+    myzh@stanford.edu
 	\and
-	Steven Tung
+	Steven Tung\\
+    steven.tung@stanford.edu
 	\and
-	Kevin Krakauer
+	Kevin Krakauer\\
+    kevinkrakauer@gmail.com
 }
@@
 
-\section*{}
+\section*{Abstract}
 We built an eventually consistent, append-only datastore focused on low
 read/write latency and high availability. In the common case, clients interact
-only with nearby nodes for extremely low latency. We do not provide overwrite
-and delete operations. This simple interface guarantees properties making client
-implementation simple: the relative order of data for a given key is consistent
-and each client communicates only with a single node.
+only with nearby nodes for extremely low latency. Data is always appended to the 
+store for a given key and the relative order of observed data is maintained.
 
@@
 system, preserving availability in the face of total and long-term partitioning.
-We believe this system will be useful in chat and distributed logging
-applications.
+We believe this system will be useful in chat, social media, and distributed
+logging applications.
 
+\vspace{-0.4cm}
 \section{Introduction}
+\vspace{-0.2cm}
 As organizations increasingly move to the cloud, applications are designed from
-the ground up with a distributed architecture (in some cases referred to as
-\textit{microservices}). Despite numerous advantages, distributed
-application design requires addressing latency and partitioning concerns that
-monolithic applications do not have (or in the case of partitioning, are not
-solvable).
+the ground up with a distributed architecture (\textit{microservices}). Despite
+numerous advantages, distributed application design requires addressing latency
+and partitioning concerns that monolithic applications do not have.
 
@@
 over other eventually consistent systems:
-
 \begin{enumerate}
 	\item Clients communicate only with their nearest follower for extremely
-		low latency.
-	\item Followers provide linearizability for client appends on a single
-		key. That is, all clients writing to the same key of the same
-		follower will see linearized writes.
+		low latency. \vspace{-0.3cm}
 	\item Eventual consistency is minimally disruptive to clients. A client
 		may read the data for key $k$ and receive data consisting of 2
-		writes' data: $d_1, d_2$. Write $x$ with data $d_x$ sent through another follower
-		preserves the ordering of $d_1$ and $d_2$, so the system
-		eventually returns $d_x, d_1, d_2$, $d_1, d_x, d_2$, or $d_1,
-		d_2, d_x$ when read.  Because of (2), writes directly to the
-		same follower are linearizable.
+		writes' data: $\{d_1, d_2\}$. Write $x$ with data $d_x$ sent
+		through another follower preserves the ordering of $d_1$ and
+		$d_2$, so the system eventually returns $\{d_x, d_1, d_2\}$,
+		$\{d_1, d_x, d_2\}$, or $\{d_1, d_2, d_x\}$ when read.
 \end{enumerate}
-
 We also gain the advantages of some more traditionally eventually consistent
-systems, such as great partition tolerance \cite{dynamo}. Together, these
+systems, such as great partition tolerance \cite{Dynamo}. Together, these
 properties are highly desirable for many append-only applications. Consider the
@@
 \begin{itemize}
-	\item \textbf{Chat} - Chat applications are inherently append-only. Users
-		in different parts of the globe expect near-instantaneous (only
-		a few seconds) latency. Users in the same building, however,
-		expect to receive messages instantaneously. Think of a global
-		company: teammates send messages and links to nearby teammates,
-		expecting them to arrive immediately. Consider a chatroom
-		occupied by a team in North American and a team in Asia. Each
-		team member sees their local peers' messages immediately and in
-		exactly the order that they are sent. It's acceptable for the a
-		message from the other team to take a bit longer and appear
-		slightly out of order in the logged chat.
-	\item \textbf{Distributed logging} - Consider a distributed application
-		running in multiple data centers on multiple continents. The
-		application is monitored in real time in each datacenter. Logs
-		are write-only, and critical for monitoring. By using the
-		append-only datastore for logging, the monitoring service:
-		\begin{itemize}
-			\item Continues operating when a datacenter is
-				partitioned from the rest of the system.
-			\item Receives logs with extremely low latency from the
-				nearby monitored application.
-			\item Has an eventually consistent log that can be
-				stored for later analysis and debugging.
-		\end{itemize}
-	% \item \textbf{Distributed data processing}
+	\item \textbf{Distributed logging} - Consider a monitoring service using
+		the append-only datastore for logs in real time. The service
+		writes and retrieves logs with extremely low latency from the
+		nearby data center. Also it is provided an eventually consistent
+		log that contains data from all of the deployments.
+	\item \textbf{Chat Application} - Consider a chatroom
+		occupied by a team in North America and a team in Asia. Each
+		team member sees their local peers' messages immediately, and
+		all messages eventually.
 \end{itemize}
 
+\section{Related Work}
+Amazon's Dynamo \cite{Dynamo} supports always-writable semantics, high partition
+tolerance, and laser-focuses on low-latency operation. Unlike our system, Dynamo
+is a key-value store. It also exposes a great deal of complexity to developers,
+who must tailor their use of Dynamo such that conflicts are resolvable and
+must manually implement conflict resolution in clients. Our system does not
+allow for conflicts. Google's GFS \cite{GFS} is also optimized for append-heavy workloads and uses a
+single master for simplicity and intelligent coordination. It supports random
+writes as well. However, it is explicitly optimized for non-latency-sensitive
+applications, and a single read can require multiple hops (the GFS master and
+chunkserver). LinkedIn's Kafka \cite{Kafka} provides eventual delivery of large
+quantities of data, but running under a publish/subscribe
+mechanism.
+
+\vspace{-0.4cm}
 \section{Design}
-% TODO: Add architecture diagram.
 Nodes in the append-only datastore are classified in a simple hierarchy as seen
-in TODO(add the architecture diagram). A single \textit{leader} directs multiple
-\textit{followers}. Clients connnect to and communicate with a nearby follower,
-likely running in the same datacenter, to minimize latency. Our design tolerates
-arbitary partitions between nodes and ensures an eventually consistent view of
+in Figure \ref{Architecture}. A single \textit{leader} directs multiple
+\textit{followers}. Clients connect to and communicate with a nearby follower,
+likely running in the same data center, to minimize latency. Our design tolerates
+arbitrary partitions between nodes and ensures an eventually consistent view of
 data.
-
-Clients are exposed the following API by followers:
+%\vspace{-0.4cm}
+% Architecture
+\begin{figure}[t]
+\includegraphics[width=8cm]{figure/SystemStructure.pdf}
+\caption{Architecture}
+\label{Architecture}
+\end{figure}
+Clients are exposed to the following API by followers:
 \begin{itemize}
 	\item \texttt{append(key, data)} - Appends \texttt{data} to the existing
-		data for \texttt{key}. Guarantees that \texttt{data} appears
-		after any data that the client already read for \texttt{key}.
-	\item \texttt{get(key)} - Gets the data for \texttt{key}, expressed as a
-		list of indexed values. For example, the returned data will be
-		the result of writes 1, 3, and 5, expressed as $\{d_1, d_3,
-		d_5\}$. Subsequent \texttt{get} requests will always return
-		$d_1$, $d_3$, and $d_5$ in the same relative order, but other
-		writes may be prepended, appended, or interspersed.
+		data for \texttt{key}. Guarantees that \texttt{data} will be
+		committed eventually if it is written to local disk.
+	\item \texttt{get(key)} - Gets the data for \texttt{key}, represented as a
+		list of values. Later calls to \texttt{get} return data in the
+		same relative order, but other data may be interleaved or
+		appended.
 \end{itemize}
+
+\subsection{Leader}
+% TODO: Explain what happens when this is partitioned. Look at gfs paper.
+The leader is responsible for globally ordering writes for each key. Like GFS's
+master \cite{GFS}, it only stores metadata rather than the value itself.
+This decreases network usage.  Also, the
+burden on disks is lighter per update, increasing writes per second and the
+average lifetime of each disk. The simplicity of a single master
+simplifies our implementation and the rest of our design, which increases
+the overall stability of the system.
+
+\subsubsection{Index number}
+The leader maintains the latest index number for all keys. Index numbers are
+globally unique and monotonically increasing. Once generated, each is mapped
+to a list of values on the requesting follower. If another follower needs that
+value, it simply needs the index number to fetch the data.
+
+\vspace{-0.4cm}
+\subsubsection{Update}
+Upon receiving an \texttt{Update} request, the leader advances the index
+number by one and updates the follower information to the incoming one.
+
+The \texttt{Leader} decides whether the requesting follower needs to sync up
+with other followers. If a mapping exists from key $k$ to index $i$, an
+\texttt{Update} carrying the same index $i$ does not need a Sync ($i$ still
+increments by 1). This may happen when 1) a key created on the leader for the
+first time, or 2) the requesting follower is the same as the recorded follower
+for the key.
+
+An \texttt{Update} to $i' < i$ indicates that the sender does not have the
+latest data. In this case the leader signals the follower to sync with the
+follower in its records that originated the append at index $i$. This may happen
+when two followers are append data to the same key concurrently. The common,
+non-concurrent case is shown in Figure \ref{CommonUpdate}.
+
+% update procedure figure
+\begin{figure}[h]
+\includegraphics[width=8cm]{figure/update.pdf}
+\caption{Update procedure}
+\label{CommonUpdate}
+\end{figure}
+
+\vspace{-0.3cm}
+\subsubsection{Broadcasting index numbers}
+The leader broadcasts ensure the eventual consistency of the system. Each
+appended value gets replicated and synced immediately between the requesting
+followers and in-record follower, but all of the other followers do not know
+it until the leader broadcasts the latest index number.
+
+Depending on the workload and use cases, we provide different ways of
+broadcasting. One is triggered per-request, which is expected to provide lower
+consistency latency but may only be used in a lower QPS scenario, since the
+broadcasting may cause a great amount of communication between the followers.
+The other one is triggered periodically, which is expected to be used in most
+cases. The leader broadcasts the updated keys and their index numbers repeatedly
+at a small, set interval.
+
+\subsubsection{Leader Fault Tolerance}
+Because the leader writes all updates to disk, it tolerates crashes and reboots.
+For greater fault tolerance, it should write to multiple disks or a remote disk
+as well.
+
+The system continues to function when the leader is partitioned, but followers
+and thus clients do not receive updates from other followers. As long as the
+partition is eventually healed, the system will propagate information correctly
+and self-heal. If for some reason there is a permanent partition, it must be
+manually worked around by changing the cluster configuration (i.e. the set of
+nodes in the system).
 
@@
 % responsible for synchronizing its the writes to it.
-The follower is our most complex component. Its responsibilities include:
+The follower is our most critical component. Its responsibilities include:
 
 \begin{itemize}
-	\item Handing \texttt{append} and \texttt{get} requests from clients.
-	\item Ordering appends.
-	\item Updating the master about appended data.
-	\item Receiving the eventual, global ordering of data from the master
-		and imposing it on data.
+	\item Handing \texttt{append} and \texttt{get} requests. \vspace{-0.4cm}
+	\item Managing data. \vspace{-0.4cm}
+	\item Updating the leader about appended data. \vspace{-0.4cm}
 	\item Syncing with other followers to communicate updates and orderings.
@@
 
-% TODO: We don't actually do the update-on-get thing yet.
-Each follower stores the entire datastore locally in memory. When clients issue
-\texttt{get(k)} requests, these are served from the in-memory datastore to minimize
-response latency. It also causes the follower to ask the leader for the most
-up-to-date data for \texttt{k}, ensuring that frequently accessed data is kept
-relatively fresh.
+\subsubsection{Data structure}
+Followers replicate the entire datastore in memory to minimize
+reading and writing latency. The data structure is described in Figure \ref{DataStructure}.
 
-\texttt{append} requests are also fast, limited by the speed of a local disk
-write. When a client appends data to key \texttt{k}, the follower writes that
-data - along with its index - to local disk before returning to the client. This
-ensures that data is not lost upon crashing and rebooting. For increased fault
-tolerance, the append should be written to multiple disks locally or to a remote
-disk. Note that throughout this section when we refer to updating local storage,
-the in-memory store is also updated immediately after data is written to disk.
+% data structure figure
+\begin{figure}[h]
+\includegraphics[width=8cm]{figure/DataStructure.pdf}
+\caption{Data procedure on Follower}
+\label{DataStructure}
+\end{figure}
 
-As soon as the append is written to disk, the follower (concurrently with
-replying to the client) sends an \texttt{update} request to the leader
-containing the key being appended to and the index that the follower has
-assigned the appended data. The leader resolves and orders the append with any
-other concurrent appends from different followers (see next section), and
-responds.
+\subsubsection{Client Interfaces}
+\textbf{Append(key, data):} Upon reviving an \texttt{Append} request, the
+follower writes the data to memory first and then writes to a local disk for data
+safety. It returns immediately after the data is written to local disk. For
+increased fault tolerance, the append should be written to multiple disks
+locally or to a remote disk. Note that throughout this section when we
+refer to updating local storage, the in-memory store is also updated immediately
+after data is written to disk.
 
-There are two types of responses. The first - and likely most common - case
-is a simple \texttt{SUCCESS} response. This indicates that the append completed
-successfully and the follower has up-to-date data for that key. The second is a
-\texttt{NEEDS\_UPDATE} response. This indicates that the data was appended, but
-another \texttt{update} preempted it. In this scenario, the follower changes the
-locally stored index of the appended data, and sends a \texttt{sync} request
-another follower indicated by the leader's response.
+The follower always appends data to the pre-update-buffer first, then issues an
+\texttt{Update} if the pre-update-buffer is not empty. If it receives an
+\texttt{Update} response with a index number, it creates a new index entry and
+moves all of the data from pre-update-buffer to the new entry.
 
-Followers use the \texttt{sync} request to get data they are missing,
-effectively ``patching the holes" in their local datastore. This ensures the
-eventually consistent property of the system.
+\noindent \textbf{Get(key):} When clients issue a \texttt{Get} request, they are
+served from the in-memory datastore. We guarantees that the data read by the
+clients are correct, but not necessarily yet globally consistent. The follower
+builds up a list with all of the data from each index as well as everything from
+the pre-update-buffer.
 
-\subsubsection{Common Case}
-When updates do not overlap, the follower handles \texttt{append} requests as
-follows:
+\subsubsection{Synchronization}
+Followers use the \texttt{Sync} request to get data they missed. This ensures
+the eventual consistency of the system. \texttt{Sync} requests also signal what
+data the sender has to the recipient, enabling the recipient to create new index
+entries in its local datastore. The \texttt{Sync} request is issued when a
+follower receives an \texttt{Update} response from the leader indicating it
+needs synchronization.
 
-\begin{enumerate}
-	\item Write the append to local storage along with the index \texttt{i}
-		the follower believes applies to the append.
-	\item Respond to the client that the request was successful.
-	\item Concurrently with (2), send an \texttt{update} to the leader that
-		an append occured on key \texttt{k} at index \texttt{i}.
-	\item The leader responds with \texttt{SUCCESS}.
-	\item The follower updates local disk to indicate that the append has
-		been commited at index \texttt{i} by the leader.
-\end{enumerate}
+Since a follower may missing multiple indices of data, a \texttt{Sync} request
+may carry a list of indices that it is asking for. However the recipient may
+also be missing some of the indices if there is a large amount of \texttt{Sync}
+requests in flight simultaneously. To mitigate this, the \texttt{Sync} response
+only replies with indices and data the recipient already has, so that the
+issuer can accept what is returned and do another \texttt{Sync} asking for
+what is left. Eventually, each follower synchronizes with and achieves global
+consistency.
 
-The data for key \texttt{k} has simply benn appended to, changing from $\{d_a,
-d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ..., d_{i-1}, d_i\}$, where $a < b <
-c < i$. Note that there may be data ``holes" between $a$, $b$, $c$, and $i$ not
-yet filled in by either a \texttt{get} or \texttt{sync}.
+Combined with the \texttt{Update} requests, Figure \ref{UpdateAndSync}
+demonstrates a message flow of synchronization in common case.
 
-\subsubsection{Conccurent Updates}
-When concurrent updates occur:
-
-% TODO: There's some ugly mixing of texttt and math notation here, particularly
-% around i'.
-\begin{enumerate}
-	\item The same as steps 1-3 above.
-	\item The leader responds with \texttt{NEEDS\_UPDATE}.
-	\item The follower updates local disk to indicate that the append has
-		been commited at index \texttt{i`} by the leader.
-	\item The follower sends a \texttt{sync} request to the follower that
-		appended the data at index \texttt{i`-1}, as indicated in the
-		leader's response.
-	\item The follower receives and stores on local disk the data it was
-		missing, including what the leader deemed to be the ``winner" of
-		index \texttt{i}.
-\end{enumerate}
-
-The data for key \texttt{k} in this case is an append and a reordering,
-changing from $\{d_a, d_b, d_c, ..., d_{i-1}\}$ to $\{d_a, d_b, d_c, ...,
-d_{i-1}, d_i, d_{i'}\}$, where $a < b < c < i$ and as above may have more holes
-to fill.
-
-% TODO: We don't yet implement the early return of data.
-It is possible for a chain of \texttt{sync} requests to form when more than 2
-followers append a key at the same time. If follower $a$ starts syncing with
-follower $b$, which is already waiting on follower $c$ to respond to a
-\texttt{sync}, $b$ will send what is knows to $a$. $a$ then retries to get
-missing data, at which point $b$ waits to respond until it has new data to send.
+% UpdateAndSync figure
+\begin{figure}[h]
+\includegraphics[width=8cm]{figure/UpdateAndSync.pdf}
+\caption{Update and Sync message flow}
+\label{UpdateAndSync}
+\end{figure}
 
 \subsubsection{Follower Fault Tolerance}
-Both \texttt{update} and \texttt{sync} requests are retried in case of network
+Both \texttt{Update} and \texttt{Sync} requests are retried in case of network
 failure. If a follower reboots during a \texttt{update}, it can inspect the disk
@@
 
-\subsection{Leader}
-% Explain what happens when this is partitioned.
-% Reference GFS and not transporting the data itself.
-It synchronizes stuff.
+\vspace{-0.4cm}
+\section{Evaluation}
+The system is written in Go and can be run manually or distributed and
+run with Docker. Nodes communicate via gRPC.
 
-\subsection{Clients}
-% Describe the API and guarantees exposed to clients.
-We have very distinguished clientele.
+We ran tests with a cluster of Google Compute Engine virtual machines. Google
+Cloud platform provides internal IP communication between VM instances, which
+means request between servers do not go through the public Internet and are thus
+fast. We deployed our system in the following locations.
+\begin{itemize}
+    \vspace{-0.4cm}
+	\item One leader in Iowa, Central USA.
+	\vspace{-0.4cm}
+	\item One follower and one client in each of the following five locations
+\vspace{-0.3cm}
+    \begin{itemize}
+    \item Iowa, central USA.
+\vspace{-0.3cm}
+	\item Los Angeles, west USA.
+\vspace{-0.3cm}
+	\item South Carolina, east USA.
+\vspace{-0.3cm}
+    \item London, west Europe.
+\vspace{-0.3cm}
+    \item Tokyo, northeast Asia.
+\vspace{-0.3cm}
+    \end{itemize}
+\end{itemize}
 
-\subsection{Corner Cases}
-% Multiple simultaneous writes from c1 and c2 to f1. How do we ensure that the
-% GL does not order w2 before w1?
-% What if the sync request can be partially satisfied. Do we not respond, or
-% respond partially and expect another sync request from the same follower?
+In the performance test, each client in the performance test sends 120,000
+\texttt{Append} against 10 keys randomly, so that in total there are
+600,000 values in the datastore globally. After all of the \texttt{Appends} are
+done, each client then retrieves the data by issuing \texttt{Get} for all of the
+10 keys every 0.5 second periodically.
 
-\section{Implementation}
+\subsection{Append Request Performance} We measures the \texttt{Append} request
+latency in Table \ref{AppendLatency}.  The result shows that \texttt{Append} is
+extremely fast, about 7.5 milliseconds on average. The 600,000 requests can be
+done in a little less than 1 second due to appending only involving
+communication with a single node.
 
-\section{Evaluation}
-We tested using some super cool gcloud stuff.
+\begin{table}[h]
+\small
+\centering 
+\begin{tabular}{ |p{1.6cm}|p{1cm}|p{0.8cm}|p{0.8cm}|p{1cm}|p{0.8cm}|  }
+\hline
+Location & S. C. & Tokyo & Iowa & London & L.A. \\
+ \hline
+ 1 Req   & 8.06 & 7.91 & 8.23 & 7.82 & 7.94\\
+\hline
+600k Req & 967.62 & 948.72 & 988.00 & 938.75 & 952.92\\
+ \hline
+\end{tabular}
+\caption{Append request latency in millisecond (S.C. for South Carolina, L.A. for Los Angeles)}
+\label{AppendLatency}
+\end{table}
 
-\subsection{Workload 1}
-Oh wow, we're so fast.
+ \vspace{-0.5cm}
+\subsection{Eventual consistency}
+One critical result is the eventual consistency, since this is what we trade off
+for the low latency read and write. The result is shown in Figure
+\ref{ConsistencyDelay}. After 0.5 seconds, all of the followers have already
+begun syncing with each other and received some of the data from each other. The
+start point is different since the clients were started sequentially. After 2.0
+seconds, the follower deployed in South Carolina finished synchronization and
+received all of the 600,000 values. After 3.5 seconds, all of the follower held
+all of the data and the system reached eventual consistency.
 
-\subsection{Workload 2}
-Oh no, we're unexpectedly not that fast.
+% Consistency delay figure
+\begin{figure}[h]
+	\subfigure[Eventual Consistency Delay]{\label{fig:c}
+		\includegraphics[width=8cm]{figure/consistency.pdf}}
+	\subfigure[Get Request Latency.]{\label{fig:g}
+		\includegraphics[width=8cm]{figure/GetRequestLatency.pdf}}
+\caption{Performance Test Result}
+\vspace{-0.4cm}
+\label{ConsistencyDelay}
+\end{figure}
 
-\section{Related Work}
-% Talk about Kafka
-This is the most filler of all filler sections. Look, a reference! \cite{dynamo}
+\subsection{Get Request Performance}
+The result of the \texttt{Get} is shown in Figure \ref{ConsistencyDelay}.
+Latency starts differently because the Follower held different amount of data.
+After 3.5 seconds, the latency of different Followers converged to between 90
+milliseconds and 100 milliseconds. Note that the implementation of \texttt{Get}
+could be optimized to achieve better performance.
 
+% Get request delay figure
+%\begin{figure}[h]
+%\includegraphics[width=8cm]{figure/GetRequestLatency.pdf}
+%\caption{Get Request Latency}
+%\label{GetRequestLatency}
+%\end{figure}
+\vspace{-0.4cm}
 \section{Future Work}
-% Use a raft group instead of a single node for each follower and GL.
-% Client fallback to other followers.
-Make it actually work.
+We could greatly increase throughput by providing a client library (rather than
+a raw GRPC interface) that is aware of the indices or latest index it holds.
+This would enable followers to selectively return only the few pieces of data a
+client is missing rather than all the data.
 
+To increase fault tolerance, we have discussed making each follower and the
+leader their own small cluster of consensus nodes. This would greatly increase
+fault tolerance and, while it may complicate the system, would not complicate
+clients or the protocols via which nodes communicate.
+
+Lastly, there may be cases where followers fail or are partitioned from their
+nearby clients. In these cases, we would like to explore whether clients can
+fall back to another follower.
+\vspace{-0.3cm}
 \section{Conclusions}
-I'd just like everyone to pat themselves on the back here.
+We built a datastore to provide low latency and high availability for append-only
+data storage. While this work is tailored to a specific set of workloads, there
+are numerous applications that can benefit from this approach.
 
+Our system can be used to support distributed services such as distributed
+monitoring and chat applications with high performance. Importantly, it exposes
+a simple interface allowing for simple clients. We believe that this makes it a
+useful tool in building distributed and microservice systems.
+\vspace{-0.4cm}
 \bibliography{paper} 

### 2020-06-11 paper/paper.tex
Finalizing paper

@@
 read/write latency and high availability. In the common case, clients interact
-only with nearby nodes for extremely low latency. Data is always appended to the 
-store for a given key and the relative order of observed data is maintained.
+only with nearby nodes for extremely low latency. Data is always append to the 
+store for a given key and the order is relatively kept.
 
@@
 system, preserving availability in the face of total and long-term partitioning.
-We believe this system will be useful in chat, social media, and distributed
-logging applications.
+We believe this system will be useful in chat, social media and distributed logging
+applications.
 
@@
 As organizations increasingly move to the cloud, applications are designed from
-the ground up with a distributed architecture (\textit{microservices}). Despite
-numerous advantages, distributed application design requires addressing latency
-and partitioning concerns that monolithic applications do not have.
+the ground up with a distributed architecture (in some cases referred to as
+\textit{microservices}). Despite numerous advantages, distributed
+application design requires addressing latency and partitioning concerns that
+monolithic applications do not have (or in the case of partitioning, are not
+solvable).
 
@@
 		low latency. \vspace{-0.3cm}
-	\item Eventual consistency is minimally disruptive to clients. A client
-		may read the data for key $k$ and receive data consisting of 2
-		writes' data: $\{d_1, d_2\}$. Write $x$ with data $d_x$ sent
-		through another follower preserves the ordering of $d_1$ and
-		$d_2$, so the system eventually returns $\{d_x, d_1, d_2\}$,
-		$\{d_1, d_x, d_2\}$, or $\{d_1, d_2, d_x\}$ when read.
+	\item Followers provide linearizability for client appends on a single
+		key. That is, all clients writing to the same key of the same
+		follower will see linearized writes.\vspace{-0.3cm}
+	\item Data and its order is eventual consistency globally
 \end{enumerate}
@@
 \begin{itemize}
-	\item \textbf{Distributed logging} - Consider a monitoring service using
-		the append-only datastore for logs in real time. The service
-		writes and retrieves logs with extremely low latency from the
-		nearby data center. Also it is provided an eventually consistent
-		log that contains data from all of the deployments.
+	\item \textbf{Distributed logging} - Consider a monitoring service using the append-only datastore records logs in real time.
+The service writes and retrieves logs with extremely low latency from the nearby data center.
+Also it is provided an eventually consistent log that has the data from all of the deployments.
 	\item \textbf{Chat Application} - Consider a chatroom
-		occupied by a team in North America and a team in Asia. Each
-		team member sees their local peers' messages immediately, and
-		all messages eventually.
+		occupied by a team in North American and a team in Asia. Each
+		team member sees their local peers' messages immediately and in
+		exactly the order that they are sent.
 \end{itemize}
@@
 is a key-value store. It also exposes a great deal of complexity to developers,
-who must tailor their use of Dynamo such that conflicts are resolvable and
+who have to tailor their use of Dynamo such that conflicts are resolvable and
 must manually implement conflict resolution in clients. Our system does not
@@
 \end{figure}
-Clients are exposed to the following API by followers:
+Clients are exposed the following API by followers:
 \begin{itemize}
 	\item \texttt{append(key, data)} - Appends \texttt{data} to the existing
-		data for \texttt{key}. Guarantees that \texttt{data} will be
-		committed eventually if it is written to local disk.
+		data for \texttt{key}. Guarantees that \texttt{data} is safe if data is written in local disk.
 	\item \texttt{get(key)} - Gets the data for \texttt{key}, represented as a
-		list of values. Later calls to \texttt{get} return data in the
-		same relative order, but other data may be interleaved or
-		appended.
+		list of values. Data written by the same client is guaranteed to be in the same order, but new data may appear later.
 \end{itemize}
@@
 \subsubsection{Index number}
-The leader maintains the latest index number for all keys. Index numbers are
-globally unique and monotonically increasing. Once generated, each is mapped
-to a list of values on the requesting follower. If another follower needs that
-value, it simply needs the index number to fetch the data.
+The Leader maintains the latest index number for all keys.
+The index numbers are globally unique and monotonically increasing.
+Once generated, it will be mapped to a list of value on the requesting follower.
+If another follower needs that value, it simply needs the index number to fetch the data.
+The index numbers are protected by mutex lock to avoid conflicting update.
 
@@
 \subsubsection{Update}
-Upon receiving an \texttt{Update} request, the leader advances the index
-number by one and updates the follower information to the incoming one.
+Upon receiving an \texttt{Update} request, the \texttt{Leader} advance the index number by one and update the follower information to the incoming one.
+Each key entry is locked before updating to avoid concurrent write issue.
 
-The \texttt{Leader} decides whether the requesting follower needs to sync up
-with other followers. If a mapping exists from key $k$ to index $i$, an
-\texttt{Update} carrying the same index $i$ does not need a Sync ($i$ still
-increments by 1). This may happen when 1) a key created on the leader for the
-first time, or 2) the requesting follower is the same as the recorded follower
-for the key.
+The \texttt{Leader} decides if the requesting follower needs to sync up with other followers.
+If a mapping exists from key $k$ to index $i$, an \texttt{Update} carrying the same index $i$ does not need an Sync (the $i$ still increments by 1).
+This may happen when 1)  a key created on the \texttt{Leader} for the first time, or 2) the requesting follower is the same as the recorded follower for the key.
 
-An \texttt{Update} to $i' < i$ indicates that the sender does not have the
-latest data. In this case the leader signals the follower to sync with the
-follower in its records that originated the append at index $i$. This may happen
-when two followers are append data to the same key concurrently. The common,
-non-concurrent case is shown in Figure \ref{CommonUpdate}.
+An \texttt{Update} to $i' < i$ indicates that the sender is not holding the latest data.
+In this case the leader signals it to sync with the follower in its records that originated the append at index $i$.
+This may happen when two followers are appending data to the same key alternatively.
+A common case is shown in Figure \ref{CommonUpdate}.
 
@@
 \subsubsection{Broadcasting index numbers}
-The leader broadcasts ensure the eventual consistency of the system. Each
-appended value gets replicated and synced immediately between the requesting
-followers and in-record follower, but all of the other followers do not know
-it until the leader broadcasts the latest index number.
+The broadcasting of the \texttt{Leader} ensures the eventual consistency of the system.
+The new appended value gets replicated and synced immediately between the requesting followers and in-record follower, but all of the other followers would not know it until the \texttt{Leader} broadcasting the latest index number.
 
-Depending on the workload and use cases, we provide different ways of
-broadcasting. One is triggered per-request, which is expected to provide lower
-consistency latency but may only be used in a lower QPS scenario, since the
-broadcasting may cause a great amount of communication between the followers.
-The other one is triggered periodically, which is expected to be used in most
-cases. The leader broadcasts the updated keys and their index numbers repeatedly
-at a small, set interval.
+Depending on the workload and use cases, we provide different ways of broadcasting.
+One is triggered per-request, which is expected to provide lower consistency latency but may only be used in a lower QPS scenario, since the broadcasting may cause a great amount of communication between the followers.
+The other one is triggered periodically, which is expected to be used in most cases.
+The \texttt{Leader} only broadcasts the updated keys and their index numbers in a short period of time.
 
@@
 \subsubsection{Client Interfaces}
-\textbf{Append(key, data):} Upon reviving an \texttt{Append} request, the
-follower writes the data to memory first and then writes to a local disk for data
-safety. It returns immediately after the data is written to local disk. For
-increased fault tolerance, the append should be written to multiple disks
-locally or to a remote disk. Note that throughout this section when we
-refer to updating local storage, the in-memory store is also updated immediately
-after data is written to disk.
+\textbf{Append(key, data):} Upon reviving \texttt{Append} request, the follower write the data to memory first and then write to a local disk for data safety.
+It returns immediately after the data is written to local disk.
+For increased fault
+tolerance, the append should be written to multiple disks locally or to a remote
+disk as well. Note that throughout this section when we refer to updating local
+storage, the in-memory store is also updated immediately after data is written
+to disk.
 
-The follower always appends data to the pre-update-buffer first, then issues an
-\texttt{Update} if the pre-update-buffer is not empty. If it receives an
-\texttt{Update} response with a index number, it creates a new index entry and
-moves all of the data from pre-update-buffer to the new entry.
+The follower always appends the data to the pre-update-buffer firstly, then issues an \texttt{Update} if the pre-update-buffer is not empty.
+If it receives an \texttt{Update} response with a index number, it creates an new index entry and moves all of the data from pre-update-buffer to the new entry.
 
-\noindent \textbf{Get(key):} When clients issue a \texttt{Get} request, they are
-served from the in-memory datastore. We guarantees that the data read by the
-clients are correct, but not necessarily yet globally consistent. The follower
-builds up a list with all of the data from each index as well as everything from
-the pre-update-buffer.
+\noindent \textbf{Get(key):} When clients issue a \texttt{Get} requests, they are served from the in-memory datastore.
+It guarantees that the data written by the local clients are up-to-date but it is not globally consistent.
+The follower builds up a list with all of the data from each index as well as everything from the pre-update-buffer.
 
 \subsubsection{Synchronization}
-Followers use the \texttt{Sync} request to get data they missed. This ensures
-the eventual consistency of the system. \texttt{Sync} requests also signal what
-data the sender has to the recipient, enabling the recipient to create new index
-entries in its local datastore. The \texttt{Sync} request is issued when a
-follower receives an \texttt{Update} response from the leader indicating it
-needs synchronization.
+Followers use the \texttt{Sync} request to get data they missed.
+This ensures the eventual consistency of the system.
+\texttt{Sync} requests also signal what data the sender has to the recipient, enabling the recipient to create new index entries in its local datastore.
+The \texttt{Sync} request is issued when a follower receives an \texttt{Update} response from the \texttt{Leader} indicating it needs synchronization.
 
-Since a follower may missing multiple indices of data, a \texttt{Sync} request
-may carry a list of indices that it is asking for. However the recipient may
-also be missing some of the indices if there is a large amount of \texttt{Sync}
-requests in flight simultaneously. To mitigate this, the \texttt{Sync} response
-only replies with indices and data the recipient already has, so that the
-issuer can accept what is returned and do another \texttt{Sync} asking for
-what is left. Eventually, each follower synchronizes with and achieves global
-consistency.
+Since a follower may missing multiple indices of data, a \texttt{Sync} request may carries a list of indices that it is asking for.
+However the recipient may also missing some of the indices if there is a large amount of \texttt{Sync} requests on the flight everywhere.
+To mitigate this, the \texttt{Sync} response only reply with indices and the data the recipient already has, so that the issuer could accept what are returned and do another \texttt{Sync} asking for what is left.
+Eventually, each follower will synchronize with  other ones and achieves the globally consistency.
 
-Combined with the \texttt{Update} requests, Figure \ref{UpdateAndSync}
-demonstrates a message flow of synchronization in common case.
+Combining with the \texttt{Update} requests, the Figure \ref{UpdateAndSync} demonstrates a message flow of synchronization in common case.
 
@@
 
-We ran tests with a cluster of Google Compute Engine virtual machines. Google
-Cloud platform provides internal IP communication between VM instances, which
-means request between servers do not go through the public Internet and are thus
-fast. We deployed our system in the following locations.
+We ran tests with a cluster of Google Compute Engine virtual
+machines.
+Google Cloud platform provides internal IP address communication between VM instances, which means the request among servers would not need to go through public network so that the speed would be much faster.
+We deployed our system in the following locations.
 \begin{itemize}
@@
 
-In the performance test, each client in the performance test sends 120,000
-\texttt{Append} against 10 keys randomly, so that in total there are
-600,000 values in the datastore globally. After all of the \texttt{Appends} are
-done, each client then retrieves the data by issuing \texttt{Get} for all of the
-10 keys every 0.5 second periodically.
+In the performance test, each client in the performance test sends 120,000 \texttt{Append} against 10 keys randomly, so that in total there would be 600,000 values in the datastore globally.
+After all of the \texttt{Appends} are done, each client then retrieves the data by issuing \texttt{Get} for all of the 10 keys every 0.5 second periodically.
 
-\subsection{Append Request Performance} We measures the \texttt{Append} request
-latency in Table \ref{AppendLatency}.  The result shows that \texttt{Append} is
-extremely fast, about 7.5 milliseconds on average. The 600,000 requests can be
-done in a little less than 1 second due to appending only involving
-communication with a single node.
+ \vspace{-0.5cm}
+\subsection{Append Request Performance}
+We measures the \texttt{Append} request latency firstly, see Table \ref{AppendLatency}.
+The result shows that the speed of \texttt{Append} is extremely fast, which is about 7.5 millisecond in average.
+And the 600,000 requests can be done in a little less than 1 second, so that the performance test can be considered as showing a QPS performance of our system.
 
@@
 \end{tabular}
-\caption{Append request latency in millisecond (S.C. for South Carolina, L.A. for Los Angeles)}
+\caption{Append request latency in millisecond(S.C. for South Carolina, L.A. for Los Angeles)}
 \label{AppendLatency}
@@
 \subsection{Eventual consistency}
-One critical result is the eventual consistency, since this is what we trade off
-for the low latency read and write. The result is shown in Figure
-\ref{ConsistencyDelay}. After 0.5 seconds, all of the followers have already
-begun syncing with each other and received some of the data from each other. The
-start point is different since the clients were started sequentially. After 2.0
-seconds, the follower deployed in South Carolina finished synchronization and
-received all of the 600,000 values. After 3.5 seconds, all of the follower held
-all of the data and the system reached eventual consistency.
+One critical result is the eventual consistency, since this is what we trade off for the low latency read and write.
+The result is shown in Figure \ref{ConsistencyDelay}.
+We can see after 0.5 second, all of the Followers have already started syncing with each other and received some of the data from each other.
+The start point is different since the clients were started not at exactly the same time.
+After 2.0 seconds, we can see that the Follower deployed in South Carolina finished synchronization and received all of the 600,000 values.
+After 3.5 seconds, all of the follower were holding all of the data and the system reached the eventual consistency.
 
@@
 
-\subsection{Get Request Performance}
-The result of the \texttt{Get} is shown in Figure \ref{ConsistencyDelay}.
-Latency starts differently because the Follower held different amount of data.
-After 3.5 seconds, the latency of different Followers converged to between 90
-milliseconds and 100 milliseconds. Note that the implementation of \texttt{Get}
-could be optimized to achieve better performance.
+\subsection{Get Request Performace}
+The result of the \texttt{Get} is shown in Figure \ref{GetRequestLatency}.
+From the figure, we can see the latency starts differently because the Follower held different amount of data.
+After 3.5 seconds, the latency of different Followers converged to between 90 milliseconds and 100 milliseconds.
+Note that the implementation of the \texttt{Get} could be optimized to achieve even better performance.
 
@@
 a raw GRPC interface) that is aware of the indices or latest index it holds.
-This would enable followers to selectively return only the few pieces of data a
-client is missing rather than all the data.
+This would enable followers to selectively return only the few missing pieces of
+data a client is missing rather than all the data.
 
@@
 We built a datastore to provide low latency and high availability for append-only
-data storage. While this work is tailored to a specific set of workloads, there
+data storage. While this work is tailed to a specific set of workloads, there
 are numerous applications that can benefit from this approach.

### 2020-06-11 paper/paper.tex
Fix format (#42)

--- 2020-06-08 paper/paper.tex
@@
 
-\vspace{-0.4cm}
+\vspace{-0.2cm}
 \subsubsection{Update}
@@
 non-concurrent case is shown in Figure \ref{CommonUpdate}.
+Since \texttt{Follower 4} only has index 5 for "key-2", the \texttt{Leader} will indicates \texttt{Follower 4} to sync up with \texttt{Follower 1} for index 6, 7, 8 ,9.
 
@@
 
-The follower always appends data to the pre-update-buffer first, then issues an
-\texttt{Update} if the pre-update-buffer is not empty. If it receives an
+The follower always appends the input data to the pre-update-buffer first, then issues an
+\texttt{Update} asynchronous, if the pre-update-buffer is not empty. When it receives an
 \texttt{Update} response with a index number, it creates a new index entry and
@@
 \section{Evaluation}
-The system is written in Go and can be run manually or distributed and
+The system is written in Golang and can be run manually or distributed and
 run with Docker. Nodes communicate via gRPC.
@@
 \begin{itemize}
-    \vspace{-0.4cm}
+    \vspace{-0.2cm}
 	\item One leader in Iowa, Central USA.
-	\vspace{-0.4cm}
+	\vspace{-0.2cm}
 	\item One follower and one client in each of the following five locations
-\vspace{-0.3cm}
+\vspace{-0.2cm}
     \begin{itemize}
     \item Iowa, central USA.
-\vspace{-0.3cm}
+\vspace{-0.2cm}
 	\item Los Angeles, west USA.
-\vspace{-0.3cm}
+\vspace{-0.2cm}
 	\item South Carolina, east USA.
-\vspace{-0.3cm}
+\vspace{-0.2cm}
     \item London, west Europe.
-\vspace{-0.3cm}
+\vspace{-0.2cm}
     \item Tokyo, northeast Asia.
-\vspace{-0.3cm}
+\vspace{-0.2cm}
     \end{itemize}
@@
 
-In the performance test, each client in the performance test sends 120,000
+In the performance test, each client sends 120,000
 \texttt{Append} against 10 keys randomly, so that in total there are
 600,000 values in the datastore globally. After all of the \texttt{Appends} are
-done, each client then retrieves the data by issuing \texttt{Get} for all of the
+done, each client then retrieves the data by issuing \texttt{Get} for the
 10 keys every 0.5 second periodically.
@@
 latency in Table \ref{AppendLatency}.  The result shows that \texttt{Append} is
-extremely fast, about 7.5 milliseconds on average. The 600,000 requests can be
+extremely fast, about 7.5 milliseconds on average. By design, the 600,000 requests are 
 done in a little less than 1 second due to appending only involving
@@
 for the low latency read and write. The result is shown in Figure
-\ref{ConsistencyDelay}. After 0.5 seconds, all of the followers have already
+\ref{fig:c}. After 0.5 seconds, all of the followers have already
 begun syncing with each other and received some of the data from each other. The
@@
 \begin{figure}[h]
-	\subfigure[Eventual Consistency Delay]{\label{fig:c}
-		\includegraphics[width=8cm]{figure/consistency.pdf}}
-	\subfigure[Get Request Latency.]{\label{fig:g}
-		\includegraphics[width=8cm]{figure/GetRequestLatency.pdf}}
+	\subfigure[Eventual Consistency Delay]{
+		\includegraphics[width=8cm]{figure/consistency.pdf}\label{fig:c}}
+	\subfigure[Get Request Latency.]{
+		\includegraphics[width=8cm]{figure/GetRequestLatency.pdf}\label{fig:g}}
 \caption{Performance Test Result}
@@
 \subsection{Get Request Performance}
-The result of the \texttt{Get} is shown in Figure \ref{ConsistencyDelay}.
+The perfomance of the \texttt{Get} is shown in Figure \ref{fig:g}.
 Latency starts differently because the Follower held different amount of data.
